---
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r load_libraries_read_data, include = FALSE}

library(ggnewscale)
library(khroma)
library(scico)
library(ggtext)

```

# Read data


```{r read_anomaly_2023_annual_mean_maps}

files <- list.files(here::here("data/"),
                    pattern = paste0({{year_anom}},"_anomaly_map_annual.csv"),
                    full.names = TRUE)

pco2_product_coarse_annual_regression <-
  read_csv(files,
           id = "product")

pco2_product_coarse_annual_regression <-
  pco2_product_coarse_annual_regression %>% 
  mutate(product = str_extract(product, "OceanSODA|SOM_FFN|CMEMS"))

```

```{r read_anomaly_2023_monthly_mean_maps}

files <- list.files(here::here("data/"),
                    pattern = paste0({{year_anom}},"_anomaly_map_monthly.csv"),
                    full.names = TRUE)

pco2_product_coarse_monthly_regression <-
  read_csv(files,
           id = "product")

pco2_product_coarse_monthly_regression <-
  pco2_product_coarse_monthly_regression %>% 
  mutate(product = str_extract(product, "OceanSODA|SOM_FFN|CMEMS"))

```

```{r read_hovmoeller_monthly_anomalies}

files <- list.files(here::here("data/"),
                    pattern = paste0({{year_anom}},"_anomaly_hovmoeller_monthly.csv"),
                    full.names = TRUE)

pco2_product_hovmoeller_monthly_regression <-
  read_csv(files,
           id = "product")

pco2_product_hovmoeller_monthly_regression <-
  pco2_product_hovmoeller_monthly_regression %>% 
  mutate(product = str_extract(product, "OceanSODA|SOM_FFN|CMEMS"))

```


```{r read_biome_annual_regression}

files <- list.files(here::here("data/"),
                    pattern = paste0({{year_anom}},"_biome_annual_regression.csv"),
                    full.names = TRUE)

pco2_product_annual_regression <-
  read_csv(files,
           id = "product")

pco2_product_annual_regression <-
  pco2_product_annual_regression %>% 
  mutate(product = str_extract(product, "OceanSODA|SOM_FFN|CMEMS"))

```

```{r read_biome_annual_detrended}

files <- list.files(here::here("data/"),
                    pattern = paste0({{year_anom}},"_biome_annual_detrended.csv"),
                    full.names = TRUE)

pco2_product_annual_detrended <-
  read_csv(files,
           id = "product")

pco2_product_annual_detrended <-
  pco2_product_annual_detrended %>% 
  mutate(product = str_extract(product, "OceanSODA|SOM_FFN|CMEMS"))

```

```{r read_biome_monthly_detrended_anomaly, eval=FALSE}

files <- list.files(here::here("data/"),
                    pattern = paste0({{year_anom}},"_biome_monthly_detrended_anomaly.csv"),
                    full.names = TRUE)

pco2_product_monthly_detrended_anomaly <-
  read_csv(files[1],
           id = "product")

pco2_product_monthly_detrended_anomaly <-
  pco2_product_monthly_detrended_anomaly %>% 
  mutate(product = str_extract(product, "OceanSODA|SOM_FFN|CMEMS"))

```



```{r read_map}

map <-
  read_rds(here::here("data/map.rds"))

key_biomes <-
  read_rds(here::here("data/key_biomes.rds"))

super_biomes <-
  read_rds(here::here("data/super_biomes.rds"))


```

# Define labels and breaks

```{r core_variables_for_analysis}

name_core <- c("fgco2", "fgco2_int", "fgco2_hov",
               "spco2", "sfco2",
               "atm_co2", "atm_fco2",
               "dco2", "dfco2",
               "wind", "kw",
               "temperature", "sol")

```

```{r order_variable_names}

# pco2_product_annual_regression %>%
#   distinct(product,name) %>%
#   print(n=50)


pco2_product_annual_detrended <- pco2_product_annual_detrended %>%
  mutate(name = factor(name, levels = name_core)) %>% 
  filter(!is.na(name))

pco2_product_annual_regression <- pco2_product_annual_regression %>%
  mutate(name = factor(name, levels = name_core)) %>% 
  filter(!is.na(name))

pco2_product_coarse_annual_regression <-
  pco2_product_coarse_annual_regression %>%
  mutate(name = factor(name, levels = name_core)) %>% 
  filter(!is.na(name))

pco2_product_coarse_monthly_regression <-
  pco2_product_coarse_monthly_regression %>%
  mutate(name = factor(name, levels = name_core)) %>% 
  filter(!is.na(name))

pco2_product_hovmoeller_monthly_regression <-
  pco2_product_hovmoeller_monthly_regression %>%
  mutate(name = factor(name, levels = name_core)) %>% 
  filter(!is.na(name))


```



```{r define_labels_and_breaks_OIA_variables_change}

labels_breaks <- function(i_name) {
  
  if (i_name == "dco2") {
    i_legend_title <- "ΔpCO<sub>2</sub><br>(µatm)"
    # i_breaks <- c(-Inf, seq(0, 80, 10), Inf)
    # i_contour_level <- 50
    # i_contour_level_abs <- 2200
  }
  
  if (i_name == "dfco2") {
    i_legend_title <- "ΔfCO<sub>2</sub><br>(µatm)"
    # i_breaks <- c(-Inf, seq(0, 80, 10), Inf)
    # i_contour_level <- 50
    # i_contour_level_abs <- 2200
  }
  
  if (i_name == "atm_co2") {
    i_legend_title <- "pCO<sub>2,atm</sub><br>(µatm)"
    # i_breaks <- c(-Inf, seq(0, 80, 10), Inf)
    # i_contour_level <- 50
    # i_contour_level_abs <- 2200
  }
  
  if (i_name == "atm_fco2") {
    i_legend_title <- "fCO<sub>2,atm</sub><br>(µatm)"
    # i_breaks <- c(-Inf, seq(0, 80, 10), Inf)
    # i_contour_level <- 50
    # i_contour_level_abs <- 2200
  }
  
  if (i_name == "sol") {
    i_legend_title <- "CO<sub>2</sub> solubility<br>(mol m<sup>-3</sup> µatm<sup>-1</sup>)"
    # i_breaks <- c(-Inf, seq(0, 80, 10), Inf)
    # i_contour_level <- 50
    # i_contour_level_abs <- 2200
  }
  
  if (i_name == "kw") {
    i_legend_title <- "K<sub>w</sub><br>(m yr<sup>-1</sup>)"
    # i_breaks <- c(-Inf, seq(0, 80, 10), Inf)
    # i_contour_level <- 50
    # i_contour_level_abs <- 2200
  }
  
  if (i_name == "spco2") {
    i_legend_title <- "pCO<sub>2,ocean</sub><br>(µatm)"
    # i_breaks <- c(-Inf, seq(0, 80, 10), Inf)
    # i_contour_level <- 50
    # i_contour_level_abs <- 2200
  }
  
  if (i_name == "sfco2") {
    i_legend_title <- "fCO<sub>2,ocean</sub><br>(µatm)"
    # i_breaks <- c(-Inf, seq(0, 80, 10), Inf)
    # i_contour_level <- 50
    # i_contour_level_abs <- 2200
  }
  
  if (i_name == "fgco2") {
    i_legend_title <- "FCO<sub>2</sub><br>(mol m<sup>-2</sup> yr<sup>-1</sup>)"
    # i_breaks <- c(-Inf, seq(0, 80, 10), Inf)
    # i_contour_level <- 50
    # i_contour_level_abs <- 2200
  }
  
  if (i_name == "fgco2_hov") {
    i_legend_title <- "FCO<sub>2</sub><br>(PgC deg<sup>-1</sup> yr<sup>-1</sup>)"
    # i_breaks <- c(-Inf, seq(0, 80, 10), Inf)
    # i_contour_level <- 50
    # i_contour_level_abs <- 2200
  }
  
  if (i_name == "fgco2_int") {
    i_legend_title <- "FCO<sub>2</sub><br>(PgC yr<sup>-1</sup>)"
    # i_breaks <- c(-Inf, seq(0, 80, 10), Inf)
    # i_contour_level <- 50
    # i_contour_level_abs <- 2200
  }
  
  if (i_name == "temperature") {
    i_legend_title <- "SST<br>(°C)"
    # i_breaks <- c(-Inf, seq(0, 80, 10), Inf)
    # i_contour_level <- 50
    # i_contour_level_abs <- 2200
  }
  
  if (i_name == "salinity") {
    i_legend_title <- "SSS"
    # i_breaks <- c(-Inf, seq(0, 80, 10), Inf)
    # i_contour_level <- 50
    # i_contour_level_abs <- 2200
  }
  
  if (i_name == "chl") {
    i_legend_title <- "lg(Chl-a)<br>(lg(mg m<sup>-3</sup>))"
    # i_breaks <- c(-Inf, seq(0, 80, 10), Inf)
    # i_contour_level <- 50
    # i_contour_level_abs <- 2200
  }
  
  if (i_name == "mld") {
    i_legend_title <- "lg(MLD)<br>(lg(m))"
    # i_breaks <- c(-Inf, seq(0, 80, 10), Inf)
    # i_contour_level <- 50
    # i_contour_level_abs <- 2200
  }
  
  if (i_name == "press") {
    i_legend_title <- "pressure<sub>atm</sub><br>(unit?)"
    # i_breaks <- c(-Inf, seq(0, 80, 10), Inf)
    # i_contour_level <- 50
    # i_contour_level_abs <- 2200
  }
  
  if (i_name == "wind") {
    i_legend_title <- "Wind <br>(m sec<sup>-1</sup>)"
    # i_breaks <- c(-Inf, seq(0, 80, 10), Inf)
    # i_contour_level <- 50
    # i_contour_level_abs <- 2200
  }
  
  all_labels_breaks <- lst(i_legend_title,
                           # i_breaks,
                           # i_contour_level,
                           # i_contour_level_abs
                           )
  
  return(all_labels_breaks)
  
}


# labels_breaks("fgco2")

x_axis_labels <-
  c(
    "dco2" = labels_breaks("dco2")$i_legend_title,
    "dfco2" = labels_breaks("dfco2")$i_legend_title,
    "atm_co2" = labels_breaks("atm_co2")$i_legend_title,
    "atm_fco2" = labels_breaks("atm_fco2")$i_legend_title,
    "sol" = labels_breaks("sol")$i_legend_title,
    "kw" = labels_breaks("kw")$i_legend_title,
    "spco2" = labels_breaks("spco2")$i_legend_title,
    "sfco2" = labels_breaks("sfco2")$i_legend_title,
    "fgco2_hov" = labels_breaks("fgco2_hov")$i_legend_title,
    "fgco2_int" = labels_breaks("fgco2_int")$i_legend_title,
    "temperature" = labels_breaks("temperature")$i_legend_title,
    "salinity" = labels_breaks("salinity")$i_legend_title,
    "chl" = labels_breaks("chl")$i_legend_title,
    "mld" = labels_breaks("mld")$i_legend_title,
    "press" = labels_breaks("press")$i_legend_title,
    "wind" = labels_breaks("wind")$i_legend_title
  )


```


# Maps

The following maps show the anomalies of each variable in `r {{year_anom}}` as provided through the pCO2 product. Anomalies are determined based on the predicted value of a linear regression model fit to the available data from 1990 to `r {{year_anom}}-1`.

Maps are first presented as annual means, and than as monthly means. Note that the 2023 predictions for the monthly maps are done individually for each month, such the mean seasonal anomaly from the annual mean is removed.

Note: The increase the computational speed, I regridded all maps to 5X5° grid.

## Annual means

### `r {{year_anom}}` anomaly

```{r anomaly_year_annual_mean_maps, fig.asp=1.5}

pco2_product_coarse_annual_regression %>%
  filter(year == {{year_anom}}) %>%
  group_split(name) %>%
  # head(1) %>%
  map(
    ~ map +
      geom_tile(data = .x,
                aes(lon, lat, fill = resid)) +
      labs(title =  paste({{year_anom}}, "anomaly")) +
      scale_fill_divergent(
        name = labels_breaks(.x %>% distinct(name))
      ) +
      theme(legend.title = element_markdown()) +
      facet_wrap(~ product, ncol = 1)
  )


```


```{r anomaly_year_annual_mean_maps_ensemble_mean, fig.asp=0.5}

pco2_product_coarse_annual_regression_ensemble <-
  pco2_product_coarse_annual_regression %>% 
  filter(year == {{year_anom}}) %>%
  group_by(name, lon, lat) %>%
  summarize(
    resid_range = max(resid) - min(resid),
    resid_mean = mean(resid),
    n = n()
  ) %>%
  ungroup() %>%
  filter(n > 1)



pco2_product_coarse_annual_regression_ensemble %>%
  mutate(product = "Ensemble mean") %>% 
  group_split(name) %>%
  # head(3) %>%
  map(
    ~ map +
      geom_tile(data = .x,
                aes(lon, lat, fill = resid_mean)) +
      geom_point(data = .x %>% filter(abs(resid_mean) < resid_range),
                aes(lon, lat, shape = "Mean < Range")) +
      labs(title =  paste({{year_anom}}, "anomaly")) +
      scale_fill_divergent(
        name = labels_breaks(.x %>% distinct(name))
      ) +
      scale_shape_manual(values = 4, name="") +
      theme(legend.title = element_markdown()) +
      facet_wrap(~ product, ncol = 1)
  )


```


```{r anomaly_year_annual_mean_maps_ensemble_mean_offset, fig.asp=1.5}

left_join(
    pco2_product_coarse_annual_regression_ensemble,
    pco2_product_coarse_annual_regression %>% filter(year == {{year_anom}})
  ) %>%
  mutate(resid_offset = resid - resid_mean) %>% 
  group_split(name) %>%
  # head(1) %>%
  map(
    ~ map +
      geom_tile(data = .x,
                aes(lon, lat, fill = resid_offset)) +
      labs(title =  paste({{year_anom}}, "offset from anomaly")) +
      scale_fill_divergent(
        name = labels_breaks(.x %>% distinct(name))
      ) +
      theme(legend.title = element_markdown()) +
      facet_wrap(~ product, ncol = 1)
  )


```


### Trends

```{r change_1990_annomaly_year_annual_mean_maps, fig.asp=1.5}

pco2_product_coarse_annual_regression %>%
  filter(name %in% name_core) %>% 
  group_by(name) %>%
  filter(year %in% c(min(year), max(year))) %>%
  ungroup() %>%
  select(-c(value, resid)) %>%
  arrange(year) %>%
  group_by(lon, lat, name, product) %>%
  mutate(change = fit - lag(fit),
         period = paste(lag(year), year, sep = "-")) %>%
  ungroup() %>%
  filter(!is.na(change)) %>% 
  group_split(name) %>%
  # head(1) %>%
  map(
    ~ map +
      geom_tile(data = .x,
                aes(lon, lat, fill = change)) +
      labs(title =  paste("Change: ",.x$period)) +
      scale_fill_divergent(name = labels_breaks(.x %>% distinct(name))) +
      theme(legend.title = element_markdown()) +
      facet_wrap(~ product, ncol = 1)
  )

```



## Monthly means

### `r {{year_anom}}` anomaly

```{r anomaly_2023_monthly_mean_maps, fig.asp=2.5}

pco2_product_coarse_monthly_regression %>%
  filter(name %in% name_core) %>%
  group_split(name) %>%
  # head(1) %>%
  map(
    ~ map +
      geom_tile(data = .x,
                aes(lon, lat, fill = resid)) +
      labs(title = paste({{year_anom}}, "anomaly")) +
      scale_fill_divergent(name = labels_breaks(.x %>% distinct(name))) +
      theme(legend.title = element_markdown()) +
      facet_grid(month ~ product)
  )

```



# Hovmoeller plots


The following Hovmoeller plots show the anomalies from the prediction of a linear/quadratic fit to the data from 1990 to `r {{year_anom}}-1`.

Hovmoeller plots are presented as monthly means. Note that the predictions for the monthly Hovmoeller plots are done individually for each month, such the mean seasonal anomaly from the annual mean is removed.


## Monthly means

### Anomalies

```{r hovmoeller_monthly_anomalies, fig.asp=0.8}


pco2_product_hovmoeller_monthly_regression %>%
  filter(name %in% name_core) %>%
  group_split(name) %>%
  # head(1) %>%
  map(
    ~ ggplot(data = .x,
             aes(decimal, lat, fill = resid)) +
      geom_raster() +
      scale_fill_divergent(name = labels_breaks(.x %>% distinct(name))) +
      theme(legend.title = element_markdown()) +
      coord_cartesian(expand = 0) +
      labs(title = "Monthly mean anomalies",
           y = "Latitude") +
      theme(axis.title.x = element_blank()) +
      facet_wrap( ~ product, ncol = 1)
  )




```

```{r hovmoeller_monthly_anomalies_ensemble_mean, fig.asp=0.4}


pco2_product_hovmoeller_monthly_regression_ensemble <-
  pco2_product_hovmoeller_monthly_regression %>% 
  group_by(name, decimal, lat) %>%
  summarize(
    resid_range = max(resid) - min(resid),
    resid_mean = mean(resid),
    n = n()
  ) %>%
  ungroup() %>%
  filter(n > 1)
  

pco2_product_hovmoeller_monthly_regression_ensemble %>%
  mutate(product = "Ensemble mean") %>% 
  group_split(name) %>%
  # head(1) %>%
  map(
    ~ ggplot(data = .x,
             aes(decimal, lat, fill = resid_mean)) +
      geom_raster() +
      scale_fill_divergent(name = labels_breaks(.x %>% distinct(name))) +
      theme(legend.title = element_markdown()) +
      coord_cartesian(expand = 0) +
      labs(title = "Monthly mean anomalies",
           y = "Latitude") +
      theme(axis.title.x = element_blank()) +
      facet_wrap( ~ product, ncol = 1)
  )


```

```{r hovmoeller_monthly_anomalies_ensemble_mean_offset, fig.asp=0.8}


left_join(
    pco2_product_hovmoeller_monthly_regression_ensemble,
    pco2_product_hovmoeller_monthly_regression
  ) %>%
  mutate(resid_offset = resid - resid_mean) %>% 
  group_split(name) %>%
  # head(1) %>%
  map(
    ~ ggplot(data = .x,
             aes(decimal, lat, fill = resid_offset)) +
      geom_raster() +
      scale_fill_divergent(name = labels_breaks(.x %>% distinct(name))) +
      theme(legend.title = element_markdown()) +
      coord_cartesian(expand = 0) +
      labs(title = "Monthly offset from ensemble mean anomalies",
           y = "Latitude") +
      theme(axis.title.x = element_blank()) +
      facet_wrap( ~ product, ncol = 1)
  )




```

# Regional means and integrals

The following plots show biome-, super biome- or global- averaged/integrated values of each variable as provided through the pCO2 product, represented here as the anomalies from the prediction of a linear/quadratic fit to the data from 1990 to `r {{year_anom}}-1`.

Anomalies are presented relative to the predicted annual mean of each year, hence preserving the seasonality.
## Anomalies

### Annual mean trends

```{r anomalies_from_annual_mean_trends_overview, fig.asp=1.5}

pco2_product_annual_detrended %>% 
  filter(biome %in% "Global",
         name %in% name_core) %>%
  ggplot(aes(month, resid, group = as.factor(year))) +
  geom_path(data = . %>% filter(!between(year, {{year_anom}}-1, {{year_anom}})),
            aes(col = year)) +
  scale_color_grayC() +
  new_scale_color() +
  geom_path(data = . %>% filter(between(year, {{year_anom}}-1, {{year_anom}})),
            aes(col = as.factor(year)),
            linewidth = 1) +
  scale_color_manual(values = c("orange", "red"),
                     guide = guide_legend(reverse = TRUE,
                                          order = 1)) +
  scale_x_continuous(breaks = seq(1, 12, 3), expand = c(0, 0)) +
  labs(title = "Anomalies from predicted annual mean | Global") +
  facet_grid(name ~ product,
             scales = "free_y",
             labeller = labeller(name = x_axis_labels),
             switch = "y"
             ) +
  theme(
    strip.text.y.left = element_markdown(),
    strip.placement = "outside",
    strip.background.y = element_blank(),
    axis.title.y = element_blank(),
    legend.title = element_blank()
  )

```


```{r anomalies_from_annual_mean_trends_key_biomes_super, fig.asp=1.5}

pco2_product_annual_detrended %>%
  filter(biome %in% super_biomes,
         name %in% name_core) %>%
  group_split(biome) %>%
  # head(1) %>%
  map(
    ~ ggplot(data = .x,
             aes(month, resid, group = as.factor(year))) +
      geom_path(data = . %>% filter(!between(year, {{year_anom}}-1, {{year_anom}})),
                aes(col = year)) +
      scale_color_grayC() +
      new_scale_color() +
      geom_path(
        data = . %>% filter(between(year, {{year_anom}} - 1, {{year_anom}})),
        aes(col = as.factor(year)),
        linewidth = 1
      ) +
      scale_color_manual(
        values = c("orange", "red"),
        guide = guide_legend(reverse = TRUE,
                             order = 1)
      ) +
      scale_x_continuous(breaks = seq(1, 12, 3), expand = c(0, 0)) +
      labs(title = paste("Anomalies from predicted annual mean |", .x$biome)) +
      facet_grid(
        name ~ product,
        scales = "free_y",
        labeller = labeller(name = x_axis_labels),
        switch = "y"
      ) +
      theme(strip.text.y.left = element_markdown(),
            strip.placement = "outside",
        strip.background.y = element_blank(),
        axis.title = element_blank(),
        legend.title = element_blank()
      )
  )


```


```{r anomalies_from_annual_mean_trends_key_biomes, fig.asp=1.5}

pco2_product_annual_detrended %>%
  filter(biome %in% key_biomes,
         name %in% name_core) %>%
  group_split(biome) %>%
  # head(1) %>%
  map(
    ~ ggplot(data = .x,
             aes(month, resid, group = as.factor(year))) +
      geom_path(data = . %>% filter(!between(year, {{year_anom}}-1, {{year_anom}})),
                aes(col = year)) +
      scale_color_grayC() +
      new_scale_color() +
      geom_path(data = . %>% filter(between(year, {{year_anom}} - 1, {{year_anom}})),
                aes(col = as.factor(year)),
                linewidth = 1) +
      scale_color_manual(
        values = c("orange", "red"),
        guide = guide_legend(reverse = TRUE,
                             order = 1)
      ) +
      scale_x_continuous(breaks = seq(1, 12, 3), expand = c(0, 0)) +
      labs(title = paste("Anomalies from predicted annual mean |", .x$biome)) +
      facet_grid(
        name ~ product,
        scales = "free_y",
        labeller = labeller(name = x_axis_labels),
        switch = "y"
      ) +
      theme(strip.text.y.left = element_markdown(),
            strip.placement = "outside",
        strip.background.y = element_blank(),
        axis.title = element_blank(),
        legend.title = element_blank()
      )
  )


```


# Flux anomaly correlation


The following plots aim to unravel the correlation between biome-, super-biome- or globally- integrated monthly flux anomalies and the corresponding anomalies of the means/integrals of each other variable.

Anomalies are first presented are first presented in absolute units. Due to the different flux magnitudes, we need to plot the globally and biome-integrated fluxes separately. Secondly, we normalize the anomalies to the monthly spread (expressed as standard deviation) of the anomalies from 1990 to 2021.

## Annual anomalies

### Absolute


```{r absolute_integrals}

pco2_product_annual_regression %>%
  filter(year == {{year_anom}},
         name %in% name_core) %>%
  mutate(region = case_when(biome == "Global" ~ "Global",
                            biome %in% super_biomes ~ "Super biomes",
                            TRUE ~ "Biomes"),
         region = factor(region, levels = c("Global", "Super biomes", "Biomes"))) %>% 
  group_split(name) %>%
  # head(1) %>%
  map(
    ~ ggplot(data = .x) +
      geom_col(aes(biome, value, fill = product),
                 position = "dodge2") +
      scale_fill_light() +
      geom_col(aes(biome, fit, group = product, col = paste0({{year_anom}},"\nlinear\nprediction")),
               position = "dodge2",
               fill = "transparent") +
      labs(y = labels_breaks(unique(.x$name))$i_legend_title,
           title = "Absolute") +
      scale_color_grey() +
      facet_grid(.~region, scales = "free_x", space = "free_x") +
      theme(legend.title = element_blank(),
            axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
            axis.title.x = element_blank(),
            axis.title.y = element_markdown(),
            strip.background = element_blank(),
            legend.position = "top")
  )
  


```


```{r absolute_anomaly_integrals}

pco2_product_annual_regression %>%
  filter(year == {{year_anom}},
         name %in% name_core) %>%
  mutate(region = case_when(biome == "Global" ~ "Global",
                            biome %in% super_biomes ~ "Super biomes",
                            TRUE ~ "Biomes"),
         region = factor(region, levels = c("Global", "Super biomes", "Biomes"))) %>% 
  group_split(name) %>%
  # head(1) %>%
  map(
    ~ ggplot(data = .x) +
      geom_col(aes(biome, value - fit, fill = product),
                 position = "dodge2") +
      scale_fill_light() +
      labs(y = labels_breaks(unique(.x$name))$i_legend_title,
           title = "Anomalies") +
      scale_color_grey() +
      facet_grid(.~region, scales = "free_x", space = "free_x") +
      theme(legend.title = element_blank(),
            axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
            axis.title.x = element_blank(),
            axis.title.y = element_markdown(),
            strip.background = element_blank(),
            legend.position = "top")
  )
  


```


```{r anomalies_correlation_absolute_annual_global, fig.asp=0.5}

pco2_product_annual_regression %>%
  filter(biome %in% "Global",
         name %in% name_core) %>%
  select(-c(value, fit)) %>%
  pivot_wider(values_from = resid) %>%
  pivot_longer(-c(product, year, biome, fgco2_int))  %>%
  group_split(name) %>%
  # head(1) %>%
  map(
    ~ ggplot(data = .x,
             aes(value, fgco2_int)) +
      geom_hline(yintercept = 0) +
      geom_point(
        data = . %>% filter(year != {{year_anom}}),
        aes(fill = year),
        shape = 21
      ) +
      geom_smooth(
        data = . %>% filter(year != {{year_anom}}),
        method = "lm",
        se = FALSE,
        fullrange = TRUE,
        aes(col = paste("Regression fit\nexcl.", {{year_anom}}))
      ) +
      scale_color_grey() +
      scale_fill_grayC() +
      new_scale_fill() +
      geom_point(
        data = . %>% filter(between(year, {{year_anom}}-1, {{year_anom}})),
        aes(fill = as.factor(year)),
        shape = 21,
        size = 2
      )  +
      scale_fill_manual(
        values = c("orange", "red"),
        guide = guide_legend(reverse = TRUE,
                             order = 1)
      ) +
      labs(y = labels_breaks("fgco2_int")$i_legend_title,
           x = labels_breaks(unique(.x$name))$i_legend_title) +
      facet_grid(
        biome ~ product,
        scales = "free_y"
      ) +
      theme(
        axis.title.x = element_markdown(),
        axis.title.y = element_markdown(),
        legend.title = element_blank()
      )
  )

```

```{r anomalies_correlation_absolute_annual_biome_super, fig.asp=1}

pco2_product_annual_regression %>%
  filter(biome %in% super_biomes,
         name %in% name_core) %>%
  select(-c(value, fit)) %>%
  pivot_wider(values_from = resid) %>%
  pivot_longer(-c(product, year, biome, fgco2_int))  %>%
  group_split(name) %>%
  # head(1) %>%
  map(
    ~ ggplot(data = .x,
             aes(value, fgco2_int)) +
      geom_hline(yintercept = 0) +
      geom_point(
        data = . %>% filter(year != {{year_anom}}),
        aes(fill = year),
        shape = 21
      ) +
      geom_smooth(
        data = . %>% filter(year != {{year_anom}}),
        method = "lm",
        se = FALSE,
        fullrange = TRUE,
        aes(col = paste("Regression fit\nexcl.", {{year_anom}}))
      ) +
      scale_color_grey() +
      scale_fill_grayC() +
      new_scale_fill() +
      geom_point(
        data = . %>% filter(between(year, {{year_anom}}-1, {{year_anom}})),
        aes(fill = as.factor(year)),
        shape = 21,
        size = 2
      )  +
      scale_fill_manual(
        values = c("orange", "red"),
        guide = guide_legend(reverse = TRUE,
                             order = 1)
      ) +
      labs(y = labels_breaks("fgco2_int")$i_legend_title,
           x = labels_breaks(unique(.x$name))$i_legend_title) +
      facet_grid(
        biome ~ product,
        scales = "free_y"
      ) +
      theme(
        axis.title.x = element_markdown(),
        axis.title.y = element_markdown(),
        legend.title = element_blank()
      )
  )

```

```{r anomalies_correlation_absolute_annual_biome_key, fig.asp=1.5}

pco2_product_annual_regression %>%
  filter(biome %in% key_biomes,
         name %in% name_core) %>%
  select(-c(value, fit)) %>%
  pivot_wider(values_from = resid) %>%
  pivot_longer(-c(product, year, biome, fgco2_int))  %>%
  group_split(name) %>%
  # head(1) %>%
  map(
    ~ ggplot(data = .x,
             aes(value, fgco2_int)) +
      geom_hline(yintercept = 0) +
      geom_point(
        data = . %>% filter(year != {{year_anom}}),
        aes(fill = year),
        shape = 21
      ) +
      geom_smooth(
        data = . %>% filter(year != {{year_anom}}),
        method = "lm",
        se = FALSE,
        fullrange = TRUE,
        aes(col = paste("Regression fit\nexcl.", {{year_anom}}))
      ) +
      scale_color_grey() +
      scale_fill_grayC() +
      new_scale_fill() +
      geom_point(
        data = . %>% filter(between(year, {{year_anom}}-1, {{year_anom}})),
        aes(fill = as.factor(year)),
        shape = 21,
        size = 2
      )  +
      scale_fill_manual(
        values = c("orange", "red"),
        guide = guide_legend(reverse = TRUE,
                             order = 1)
      ) +
      labs(y = labels_breaks("fgco2_int")$i_legend_title,
           x = labels_breaks(unique(.x$name))$i_legend_title) +
      facet_grid(
        biome ~ product,
        scales = "free_y"
      ) +
      theme(
        axis.title.x = element_markdown(),
        axis.title.y = element_markdown(),
        legend.title = element_blank()
      )
  )

```

## Monthly anomalies

### Absolute

```{r anomalies_correlation_absolute_monthly_biome_key, fig.asp=1.5, eval=FALSE}

pco2_product_monthly_detrended_anomaly %>%
  filter(biome %in% super_biomes) %>%
  pivot_longer(-c(year, month, biome, fgco2_int))  %>%
  group_split(name) %>%
  head(1) %>%
  map(
    ~ ggplot(data = .x,
             aes(value, fgco2_int)) +
      geom_hline(yintercept = 0) +
      geom_point(
        data = . %>% filter(year < {{year_anom}}),
        aes(col = paste(min(year), max(year), sep = "-")),
        alpha = 0.2
      ) +
      geom_smooth(
        data = . %>% filter(year < {{year_anom}}),
        aes(col = paste(min(year), max(year), sep = "-")),
        method = "lm",
        se = FALSE,
        fullrange = TRUE
      )  +
      scale_color_grey(name = "") +
      new_scale_color() +
      geom_path(data = . %>% filter(between(year, {{year_anom}}-1, {{year_anom}})),
                aes(col = as.factor(month), group = 1))  +
      geom_point(
        data = . %>% filter(between(year, {{year_anom}}-1, {{year_anom}})),
        aes(fill = as.factor(month)),
        shape = 21,
        size = 3
      )  +
      scale_color_scico_d(
        palette = "buda",
        guide = guide_legend(reverse = TRUE,
                             order = 1),
        name = "Month\nof 2023"
      ) +
      scale_fill_scico_d(
        palette = "buda",
        guide = guide_legend(reverse = TRUE,
                             order = 1),
        name = "Month\nof 2023"
      ) +
      facet_wrap( ~ biome, ncol = 3, scales = "free_x") +
      labs(
        title = "Super biome integrated fluxes",
        y = labels_breaks("fgco2_int")$i_legend_title,
        x = labels_breaks(.x %>% distinct(name))$i_legend_title
      ) +
      theme(axis.title.x = element_markdown(),
            axis.title.y = element_markdown())
  )

```



