# {{year_anom}} anomalies

# Functions

## Anomaly detection

For the detection of anomalies at any point in time and space, we fit regression models and compare the fitted to the actual value.

We use linear regression models for all parameters, except for `r print(name_quadratic_fit)`, which are approximated with quadratic fits.

The regression models are fitted to all data since `r print(paste(start_year))`, except {{year_anom}}.


```{r anomaly_detection_anomaly}

anomaly_determination <- function(df,...) {
  
  group_by <- quos(...)
  # group_by <- quos(lon, lat)
  # df <- pco2_product_map_annual
  
  # Linear regression models

  df_lm <-
    df %>%
    filter(year != {{year_anom}},
           !(name %in% name_quadratic_fit)) %>%
    drop_na() %>%
    nest(data = -c(name, !!!group_by)) %>%
    mutate(fit = map(data, ~ flm(
      formula = value ~ year, data = .x
    )))
  
  df_lm <-
    left_join(
      df_lm %>%
        unnest_wider(fit) %>%
        select(name, !!!group_by,
               intercept = `(Intercept)`,  slope = year) %>%
        mutate(intercept = as.vector(intercept),
               slope = as.vector(slope)),
      df
    ) %>%
    mutate(fit = intercept + year * slope) %>%
    select(name, !!!group_by, year, fit, value) %>%
    mutate(resid = value - fit)

  # df_lm <-
  #   df %>%
  #   filter(year != {{year_anom}},
  #          !(name %in% name_quadratic_fit)) %>%
  #   drop_na() %>% 
  #   nest(data = -c(name, !!!group_by)) %>%
  #   mutate(
  #     fit = map(data, ~ lm(value ~ year, data = .x)),
  #     tidied = map(fit, tidy),
  #     augmented = map(fit, augment)
  #   )
  # 
  # 
  # df_lm_year_anom <-
  #   full_join(
  #     df_lm %>%
  #       unnest(tidied) %>%
  #       select(name, !!!group_by, term, estimate) %>%
  #       pivot_wider(names_from = term,
  #                   values_from = estimate) %>%
  #       mutate(fit = `(Intercept)` + year * {{year_anom}}) %>%
  #       select(name, !!!group_by, fit) %>%
  #       mutate(year = {{year_anom}}),
  #     df %>%
  #       filter(year == {{year_anom}},
  #              !(name %in% name_quadratic_fit))
  #   ) %>%
  #   mutate(resid = value - fit)
  # 
  # 
  # df_lm <-
  #   bind_rows(
  #     df_lm %>%
  #       unnest(augmented) %>%
  #       select(name, !!!group_by, year, value, fit = .fitted, resid = .resid),
  #     df_lm_year_anom
  #   )
  # 
  # rm(df_lm_year_anom)
  
  # Quadratic regression models
  
  if(any(df %>% distinct(name) %>% pull() %in% name_quadratic_fit)){
  
  df_quadratic <-
    df %>%
    filter(year != {{year_anom}},
           name %in% name_quadratic_fit) %>%
    drop_na() %>% 
    nest(data = -c(name, !!!group_by)) %>%
    mutate(
      fit = map(data, ~ flm(
        formula = value ~ year + I(year ^ 2), data = .x))
    )
  
  df_quadratic <-
    left_join(
      df_quadratic %>%
        unnest_wider(fit) %>%
        select(name, !!!group_by,
               intercept = `(Intercept)`, slope = year, slope_squared = `I(year^2)`) %>%
        mutate(intercept = as.vector(intercept),
               slope = as.vector(slope),
               slope_squared = as.vector(slope_squared)),
      df
    ) %>%
    mutate(fit = intercept + year * slope + year^2 * slope_squared) %>%
    select(name, !!!group_by, year, fit, value) %>%
    mutate(resid = value - fit)
  
  
  # df_quadratic <-
  #   df %>%
  #   filter(year != {{year_anom}},
  #          name %in% name_quadratic_fit) %>%
  #   nest(data = -c(name, !!!group_by)) %>%
  #   mutate(
  #     fit = map(data, ~ lm(value ~ year + I(year ^ 2), data = .x)),
  #     tidied = map(fit, tidy),
  #     augmented = map(fit, augment)
  #   )
  # 
  # df_quadratic_year_anom <-
  #   full_join(
  #     df_quadratic %>%
  #       unnest(tidied) %>%
  #       select(name, !!!group_by, term, estimate) %>%
  #       pivot_wider(names_from = term,
  #                   values_from = estimate) %>%
  #       mutate(fit = `(Intercept)` + year * {{year_anom}} + `I(year^2)` * {{year_anom}} ^ 2) %>%
  #       select(name, !!!group_by, fit) %>%
  #       mutate(year = {{year_anom}}),
  #     df %>%
  #       filter(year == {{year_anom}},
  #              name %in% name_quadratic_fit)
  #   ) %>%
  #   mutate(resid = value - fit)
  # 
  # 
  # df_quadratic <-
  #   bind_rows(
  #     df_quadratic %>%
  #       unnest(augmented) %>%
  #       select(name, !!!group_by, year, value, fit = .fitted, resid = .resid),
  #     df_quadratic_year_anom
  #   )
  # 
  # rm(df_quadratic_year_anom)
  
  # Join linear and quadratic regression results
  
  df_anomaly <-
    bind_rows(df_lm,
              df_quadratic)
  
  rm(df_lm,
     df_quadratic)
  
  } else{
    
    df_anomaly <- df_lm
    
    rm(df_lm)
  }
  
  df_anomaly <-
    df_anomaly %>%
    arrange(year)
  
  
  return(df_anomaly)
  
}

```

## Seasonality plots

```{r seasonality_plot_function}

warm_color <- "#B84A60FF"
cold_color <- "#16877CFF"

p_season <- function(df, 
                     dim_row = "name", 
                     dim_col = "biome", 
                     title = NULL, 
                     var = "resid",
                     scales = "free_y") {
  
  p <- ggplot(data = df,
              aes(month, !!ensym(var)))
  
  if(var == "resid"){
      p <- p +
        geom_hline(yintercept = 0, linewidth =0.5)
    
  }
  
  
  
  p <- p +
      geom_path(data = . %>% filter(year != {{year_anom}}),
                aes(group = as.factor(year),
                    col = as.factor(paste(min(year), max(year), sep = "-"))), 
                alpha = 0.5)+
      geom_path(data = . %>% 
                  filter(year != {{year_anom}}) %>% 
                  group_by_at(vars(month, dim_col, dim_row)) %>% 
                  summarise(!!ensym(var) := mean(!!ensym(var))),
                aes(col = "Climatological\nmean"), 
                linewidth = 1) +
    scale_color_manual(values = c("grey", "black"),
                       guide = guide_legend(order = 2,
                                            reverse = TRUE)) +
    new_scale_color()+
    geom_path(data = . %>% filter(year == {{year_anom}}),
                aes(col = as.factor(year)),
                linewidth = 1) +
      scale_color_manual(
        values = warm_color,
        guide = guide_legend(order = 1)
      ) +
      scale_x_continuous(breaks = seq(1, 12, 3), expand = c(0, 0)) +
      labs(title = title,
           x = "Month")
  
    if(df %>% filter(name == "fgco2") %>% nrow() > 0 & "value" %in% names(df)){
    
    df_sink <- df %>% 
      filter(year == {{year_anom}},
             name == "fgco2")
    
      p <- p +
          geom_point(data = df_sink %>% filter(value < 0),
             aes(shape = "Sink"), fill = "white") +
          geom_point(data = df_sink %>% filter(value >= 0),
             aes(shape = "Source"), fill = "white") +
        scale_shape_manual(values = c(25,24))
    
  }
  
  
  if (!(is.null(dim_col))) {
    p <- p +
      facet_grid(
        as.formula(paste(dim_row, "~", dim_col)),
        scales = scales,
        labeller = labeller(name = x_axis_labels),
        switch = "y"
      )
    
    
  } else {
    p <- p +
      facet_grid(
        as.formula(paste(dim_row, "~ .")),
        scales = "free_y",
        labeller = labeller(name = x_axis_labels),
        switch = "y"
      )
  }
  
  p <- p +
    theme(
      strip.text.y.left = element_markdown(),
      strip.placement = "outside",
      strip.background.y = element_blank(),
      axis.title.y = element_blank(),
      legend.title = element_blank()
    )
  
  p
  
}
```

## fCO2 decomposition

```{r fCO2_decomposition_function}

fco2_decomposition <- function(df, ...) {
  
  group_by <- quos(...)
  # group_by <- quos(lon, lat, month)
  # group_by <- quos(biome, year, month)
  
  pco2_product_biome_monthly_fCO2_decomposition <-
    df %>%
    filter(name %in% c("temperature", "sfco2"))
  
  pco2_product_biome_monthly_fCO2_decomposition <-
    inner_join(
      pco2_product_biome_monthly_fCO2_decomposition %>%
        filter(name == "temperature") %>%
        select(-c(value, fit)) %>%
        pivot_wider(values_from = resid),
      pco2_product_biome_monthly_fCO2_decomposition %>%
        filter(name == "sfco2") %>%
        select(-c(value, resid)) %>%
        pivot_wider(values_from = fit)
    )
  
  pco2_product_biome_monthly_fCO2_decomposition <-
    pco2_product_biome_monthly_fCO2_decomposition %>%
    mutate(sfco2_therm = (sfco2 * exp(0.0423 * temperature)) - sfco2)
  
  
  pco2_product_biome_monthly_fCO2_decomposition <-
    inner_join(
      pco2_product_biome_monthly_fCO2_decomposition,
      df %>%
        filter(name %in% c("sfco2")) %>%
        select(-c(value, fit, name)) %>%
        rename(sfco2_total = resid)
    )
  
  
  pco2_product_biome_monthly_fCO2_decomposition <-
    pco2_product_biome_monthly_fCO2_decomposition %>%
    mutate(sfco2_nontherm = sfco2_total - sfco2_therm)
  
  pco2_product_biome_monthly_fCO2_decomposition <-
    pco2_product_biome_monthly_fCO2_decomposition %>%
    select(-c(temperature, sfco2)) %>%
    pivot_longer(starts_with("sfco2"),
                 values_to = "resid")
  
}


```

## Flux attribution

```{r flux_attribution_function}

flux_attribution <- function(df, ...) {
  
  group_by <- quos(...)
  # group_by <- quos(lon, lat, month)
  
  pco2_product_flux_attribution <-
    df %>%
    filter(name %in% c("dfco2", "kw_sol", "fgco2"))
  
  
  pco2_product_flux_attribution <-
    inner_join(
      pco2_product_flux_attribution %>%
        select(-c(value, fit)) %>%
        pivot_wider(values_from = resid,
                    names_prefix = "resid_"),
      pco2_product_flux_attribution %>%
        select(-c(value, resid)) %>%
        filter(name != "fgco2") %>%
        pivot_wider(values_from = fit)
    )
  
    pco2_product_flux_attribution <-
    pco2_product_flux_attribution %>%
    mutate(
      resid_fgco2_dfco2 = resid_dfco2 * kw_sol,
      resid_fgco2_kw_sol = resid_kw_sol * dfco2,
      resid_fgco2_dfco2_kw_sol = resid_dfco2 * resid_kw_sol
      # resid_fgco2_sum = resid_fgco2_dfco2 + resid_fgco2_kw_sol + resid_fgco2_dfco2_kw_sol
    )
  
  # pco2_product_flux_attribution <-
  #   pco2_product_flux_attribution %>%
  #   mutate(resid_fgco2_offset = resid_fgco2 - resid_fgco2_sum)
  
  pco2_product_flux_attribution <-
    pco2_product_flux_attribution %>%
    select(!!!group_by, starts_with("resid_fgco2")) %>%
    pivot_longer(starts_with("resid_"),
                 values_to = "resid")
  
  
  pco2_product_flux_attribution <-
    pco2_product_flux_attribution %>%
    filter(str_detect(name, "dfco2|kw_sol")) %>% 
    mutate(name = factor(
      name,
      levels = c(
        "resid_fgco2",
        "resid_fgco2_dfco2",
        "resid_fgco2_kw_sol",
        "resid_fgco2_dfco2_kw_sol",
        "resid_fgco2_sum",
        "resid_fgco2_offset"
      )
    ))
  
}

```

# Maps

The following maps show the absolute state of each variable in {{year_anom}} as provided through the pCO2 product, the change in that variable from 1990 to {{year_anom}}, as well es the anomalies in {{year_anom}}. Changes and anomalies are determined based on the predicted value of a linear regression model fit to the data from 1990 to `r {{year_anom}}-1`.

Maps are first presented as annual means, and than as monthly means. Note that the {{year_anom}} predictions for the monthly maps are done individually for each month, such the mean seasonal anomaly from the annual mean is removed.

Note: The increase the computational speed, I regridded all maps to 5X5° grid.

## Annual means

### {{year_anom}} absolute

```{r {{year_anom}}_map_annual_absolute, fig.asp=0.6, results = FALSE}

pco2_product_map_annual_anomaly <-
  pco2_product_map_annual %>%
  drop_na() %>% 
  anomaly_determination(lon, lat)

pco2_product_map_annual_anomaly <-
  pco2_product_map_annual_anomaly %>%
  drop_na()

pco2_product_map_annual_anomaly %>%
  filter(year == {{year_anom}},
         !(name %in% name_divergent)) %>%
  group_split(name) %>%
  head(1) %>%
  map(
    ~ map +
      geom_tile(data = .x,
                aes(lon, lat, fill = value)) +
      labs(title = paste("Annual mean", {{year_anom}})) +
      scale_fill_viridis_c(name = labels_breaks(.x %>% distinct(name))) +
      guides(
        fill = guide_colorbar(
          barheight = unit(0.3, "cm"),
          barwidth = unit(6, "cm"),
          ticks = TRUE,
          ticks.colour = "grey20",
          frame.colour = "grey20",
          label.position = "top",
          direction = "horizontal"
        )
      ) +
      theme(legend.title = element_markdown(),
            legend.position = "top")
  )

pco2_product_map_annual_anomaly %>%
  filter(year == {{year_anom}},
         name %in% name_divergent) %>%
  group_split(name) %>% 
  head(1) %>%
  map( ~ map +
         geom_tile(data = .x,
                   aes(lon, lat, fill = value)) +
         labs(title = paste("Annual mean", {{year_anom}})) +
         scale_fill_gradientn(
           colours = cmocean("curl")(100),
           rescaler = ~ scales::rescale_mid(.x, mid = 0),
           name = labels_breaks(.x %>% distinct(name)),
           limits = c(quantile(.x$value, .01), quantile(.x$value, .99)),
           oob = squish
         ) +
      guides(
        fill = guide_colorbar(
          barheight = unit(0.3, "cm"),
          barwidth = unit(6, "cm"),
          ticks = TRUE,
          ticks.colour = "grey20",
          frame.colour = "grey20",
          label.position = "top",
          direction = "horizontal"
        )
      ) +
      theme(legend.title = element_markdown(),
            legend.position = "top")
  )
  
```

### Trends

```{r {{year_anom}}_map_annual_change, fig.asp=0.6, results = FALSE}

pco2_product_map_annual_anomaly %>%
  group_by(name) %>%
  filter(year %in% c(min(year), max(year), {{year_anom}})) %>%
  ungroup() %>%
  select(-c(value, resid)) %>%
  filter(year %in% c(min(year), max(year))) %>%
  arrange(year) %>% 
  group_by(lon, lat, name) %>%
  mutate(change = fit - lag(fit),
         period = paste(lag(year), year, sep = "-")) %>%
  ungroup() %>%
  filter(!is.na(change)) %>%
  group_split(name) %>% 
  head(1) %>%
  map(
    ~ map +
      geom_tile(data = .x,
                aes(lon, lat, fill = change)) +
      labs(title =  paste("Change: ",.x$period)) +
      scale_fill_gradientn(
        colours = cmocean("curl")(100),
        rescaler = ~ scales::rescale_mid(.x, mid = 0),
        name = labels_breaks(.x %>% distinct(name)),
        limits = c(quantile(.x$change, .01), quantile(.x$change, .99)),
        oob = squish
      ) +
      guides(
        fill = guide_colorbar(
          barheight = unit(0.3, "cm"),
          barwidth = unit(6, "cm"),
          ticks = TRUE,
          ticks.colour = "grey20",
          frame.colour = "grey20",
          label.position = "top",
          direction = "horizontal"
        )
      ) +
      theme(legend.title = element_markdown(),
            legend.position = "top")
  )

```

### {{year_anom}} anomaly

```{r {{year_anom}}_map_annual_anomaly, fig.asp=0.6, results = FALSE}

pco2_product_map_annual_anomaly %>%
  filter(year == {{year_anom}}) %>%
  group_split(name) %>% 
  head(1) %>%
  map( ~ map +
         geom_tile(data = .x,
                   aes(lon, lat, fill = resid)) +
         labs(title =  paste({{year_anom}},"anomaly")) +
         scale_fill_gradientn(
           colours = cmocean("curl")(100),
           rescaler = ~ scales::rescale_mid(.x, mid = 0),
           name = labels_breaks(.x %>% distinct(name)),
           limits = c(quantile(.x$resid, .01), quantile(.x$resid, .99)),
           oob = squish
         )+
      guides(
        fill = guide_colorbar(
          barheight = unit(0.3, "cm"),
          barwidth = unit(6, "cm"),
          ticks = TRUE,
          ticks.colour = "grey20",
          frame.colour = "grey20",
          label.position = "top",
          direction = "horizontal"
        )
      ) +
      theme(legend.title = element_markdown(),
            legend.position = "top")
  )

```

### SST flux slope

```{r map_annual_anomaly_slope, fig.asp=0.6}

pco2_product_map_annual_anomaly_slope <-
pco2_product_map_annual_anomaly %>%
  filter(year != {{year_anom}}) %>% 
  select(year, lon, lat, resid, name) %>% 
  pivot_wider(values_from = resid) %>%
  select(lon, lat, fgco2, temperature) %>%
  drop_na() %>% 
  nest(data = -c(lon, lat)) %>%
  mutate(fit = map(data, ~ flm(
    formula = fgco2 ~ temperature, data = .x
  )))
  
pco2_product_map_annual_anomaly_slope <-
  pco2_product_map_annual_anomaly_slope %>%
  unnest_wider(fit) %>%
  select(lon, lat, slope = temperature) %>%
  mutate(slope = as.vector(slope))

map +
  geom_tile(data = pco2_product_map_annual_anomaly_slope, 
            aes(lon, lat, fill = slope)) +
  scale_fill_gradientn(
    colours = cmocean("curl")(100),
    rescaler = ~ scales::rescale_mid(.x, mid = 0),
    limits = c(
      quantile(pco2_product_map_annual_anomaly_slope$slope,.01),
      quantile(pco2_product_map_annual_anomaly_slope$slope, .99)),
    oob = squish
  ) +
  labs(title = "Correlation of historic annual flux and SST anomalies") +
  guides(
    fill = guide_colorbar(
      barheight = unit(0.3, "cm"),
      barwidth = unit(6, "cm"),
      ticks = TRUE,
      ticks.colour = "grey20",
      frame.colour = "grey20",
      label.position = "top",
      direction = "horizontal"
    )
  ) +
  theme(legend.title = element_markdown(), legend.position = "top")

```



```{r write_anomaly_anomaly_year_annual_mean_maps, results = FALSE}

pco2_product_map_annual_anomaly_slope %>%
  write_csv(
    paste0(
      "../data/",
      "{{product_name}}",
      "_",
      "{{year_anom}}",
      "_map_annual_anomaly_slope.csv"
    )
  )

pco2_product_map_annual_anomaly %>%
  filter(year == {{year_anom}}) %>%
  write_csv(
    paste0(
      "../data/",
      "{{product_name}}",
      "_",
      "{{year_anom}}",
      "_map_annual_anomaly.csv"
    )
  )

rm(pco2_product_map_annual_anomaly,
   pco2_product_map_annual_anomaly_slope)
gc()

```




## Monthly means

### {{year_anom}} absolute

```{r {{year_anom}}_map_monthly_absolute, fig.asp=1.6, results = FALSE}

pco2_product_map_monthly_anomaly <-
  pco2_product_map_monthly %>%
  drop_na() %>% 
  anomaly_determination(lon, lat, month)

pco2_product_map_monthly_anomaly <-
  pco2_product_map_monthly_anomaly %>% 
  drop_na()



pco2_product_map_monthly_anomaly %>%
  filter(year == {{year_anom}}, !(name %in% name_divergent)) %>%
  group_split(name) %>%
  head(1) %>%
  map(
    ~ map +
      geom_tile(data = .x, aes(lon, lat, fill = value)) +
      labs(title = paste("Monthly means", {{year_anom}})) +
      scale_fill_viridis_c(name = labels_breaks(.x %>% distinct(name))) +
      guides(
        fill = guide_colorbar(
          barheight = unit(0.3, "cm"),
          barwidth = unit(6, "cm"),
          ticks = TRUE,
          ticks.colour = "grey20",
          frame.colour = "grey20",
          label.position = "top",
          direction = "horizontal"
        )
      ) +
      theme(legend.title = element_markdown(), legend.position = "top") +
      facet_wrap(~ month, ncol = 2)
  )

pco2_product_map_monthly_anomaly %>%
  filter(year == {{year_anom}}, name %in% name_divergent) %>%
  group_split(name) %>%
  head(1) %>%
  map(
    ~ map +
      geom_tile(data = .x, aes(lon, lat, fill = value)) +
      labs(title = paste("Monthly means", {{year_anom}})) +
      scale_fill_gradientn(
        colours = cmocean("curl")(100),
        rescaler = ~ scales::rescale_mid(.x, mid = 0),
        name = labels_breaks(.x %>% distinct(name)),
        limits = c(quantile(.x$value, .01), quantile(.x$value, .99)),
        oob = squish
      ) +
      guides(
        fill = guide_colorbar(
          barheight = unit(0.3, "cm"),
          barwidth = unit(6, "cm"),
          ticks = TRUE,
          ticks.colour = "grey20",
          frame.colour = "grey20",
          label.position = "top",
          direction = "horizontal"
        )
      ) +
      theme(legend.title = element_markdown(), legend.position = "top") +
      facet_wrap( ~ month, ncol = 2)
  )

```

### Trends

```{r {{year_anom}}_map_monthly_change, fig.asp=1.6, results = FALSE, eval=FALSE}

pco2_product_map_monthly_anomaly %>%
  group_by(name) %>%
  filter(year %in% c(min(year), max(year))) %>%
  ungroup() %>%
  select(-c(value, resid)) %>%
  arrange(year) %>%
  group_by(lon, lat, name, month) %>%
  mutate(change = fit - lag(fit),
         period = paste(lag(year), year, sep = "-")) %>%
  ungroup() %>%
  filter(!is.na(change)) %>%
  group_split(name) %>%
  head(1) %>%
  map(
    ~ map +
      geom_tile(data = .x, aes(lon, lat, fill = change)) +
      labs(title =  paste("Change: ", .x$period)) +
      scale_fill_gradientn(
        colours = cmocean("curl")(100),
        rescaler = ~ scales::rescale_mid(.x, mid = 0),
        name = labels_breaks(.x %>% distinct(name)),
        limits = c(quantile(.x$change, .01), quantile(.x$change, .99)),
        oob = squish
      ) +
      guides(
        fill = guide_colorbar(
          barheight = unit(0.3, "cm"),
          barwidth = unit(6, "cm"),
          ticks = TRUE,
          ticks.colour = "grey20",
          frame.colour = "grey20",
          label.position = "top",
          direction = "horizontal"
        )
      ) +
      theme(legend.title = element_markdown(), legend.position = "top") +
      facet_wrap(~ month, ncol = 2)
  )

```

### {{year_anom}} anomaly

```{r {{year_anom}}_map_monthly_anomaly, fig.asp=1.6, results = FALSE}

pco2_product_map_monthly_anomaly %>%
  filter(year == {{year_anom}}) %>%
  group_split(name) %>%
  head(1) %>%
  map(
    ~ map +
      geom_tile(data = .x, aes(lon, lat, fill = resid)) +
      labs(title = paste({{year_anom}}, "anomaly")) +
      scale_fill_gradientn(
        colours = cmocean("curl")(100),
        rescaler = ~ scales::rescale_mid(.x, mid = 0),
        name = labels_breaks(.x %>% distinct(name)),
        limits = c(quantile(.x$resid, .01), quantile(.x$resid, .99)),
        oob = squish
      ) +
      guides(
        fill = guide_colorbar(
          barheight = unit(0.3, "cm"),
          barwidth = unit(6, "cm"),
          ticks = TRUE,
          ticks.colour = "grey20",
          frame.colour = "grey20",
          label.position = "top",
          direction = "horizontal"
        )
      ) +
      theme(legend.title = element_markdown(), legend.position = "top") +
      facet_wrap( ~ month, ncol = 2)
  )

```


### fCO2 decomposition

```{r fCO2_decomposition_monthly_maps, fig.asp=1.8}

pco2_product_map_monthly_fCO2_decomposition <-
  fco2_decomposition(pco2_product_map_monthly_anomaly,
                     year, month, lon, lat)


# pco2_product_map_monthly_fCO2_decomposition %>%
#   filter(year == {{year_anom}}) %>% 
#   mutate(product == {{product_name}}) %>% 
#   group_split(product) %>%
#   head(1) %>%
#   map(
#     ~ map +
#       geom_tile(data = .x,
#                 aes(lon, lat, fill = resid)) +
#       labs(title = .x$product) +
#       scale_fill_gradientn(
#         colours = cmocean("curl")(100),
#         rescaler = ~ scales::rescale_mid(.x, mid = 0),
#         name = labels_breaks("sfco2"),
#         limits = c(quantile(.x$resid, .01), quantile(.x$resid, .99)),
#         oob = squish
#       ) +
#       facet_grid(month ~ name,
#                  labeller = labeller(name = x_axis_labels)) +
#       guides(
#         fill = guide_colorbar(
#           barheight = unit(0.3, "cm"),
#           barwidth = unit(6, "cm"),
#           ticks = TRUE,
#           ticks.colour = "grey20",
#           frame.colour = "grey20",
#           label.position = "top",
#           direction = "horizontal"
#         )
#       ) +
#       theme(legend.title = element_markdown(),
#             legend.position = "top")
#   )
```

```{r fCO2_decomposition_annual_maps, fig.asp=1.5}

pco2_product_map_annual_fCO2_decomposition <-
  pco2_product_map_monthly_fCO2_decomposition %>% 
  select(year, lat, lon, name, resid) %>% 
  fgroup_by(year, lat, lon, name) %>% 
  fmean()

gc()

map +
  geom_tile(data = pco2_product_map_annual_fCO2_decomposition %>%
              filter(year == {{year_anom}}), aes(lon, lat, fill = resid)) +
  scale_fill_gradientn(
    colours = cmocean("curl")(100),
    rescaler = ~ scales::rescale_mid(.x, mid = 0),
    name = labels_breaks("sfco2"),
    limits = c(
      quantile(pco2_product_map_annual_fCO2_decomposition$resid, .01),
      quantile(pco2_product_map_annual_fCO2_decomposition$resid, .99)
    ),
    oob = squish
  ) +
  facet_wrap( ~ name,
              ncol = 2,
              labeller = labeller(name = x_axis_labels)) +
  guides(
    fill = guide_colorbar(
      barheight = unit(0.3, "cm"),
      barwidth = unit(6, "cm"),
      ticks = TRUE,
      ticks.colour = "grey20",
      frame.colour = "grey20",
      label.position = "top",
      direction = "horizontal"
    )
  ) +
  theme(legend.title = element_markdown(), legend.position = "top")


```

### Flux attribution

```{r flux_attribution_seasonal_maps, fig.asp=1.8}

pco2_product_map_monthly_flux_attribution <-
  flux_attribution(pco2_product_map_monthly_anomaly,
                   year, month, lon, lat)

# pco2_product_map_monthly_flux_attribution %>%
#   filter(year == {{year_anom}}) %>% 
#   drop_na() %>% 
#   mutate(product == {{product_name}}) %>% 
#   group_split(product) %>%
#   head(1) %>%
#   map(
#     ~ map +
#       geom_tile(data = .x,
#                 aes(lon, lat, fill = resid)) +
#       labs(subtitle = .x$product) +
#       scale_fill_gradientn(
#         colours = cmocean("curl")(100),
#         rescaler = ~ scales::rescale_mid(.x, mid = 0),
#         name = labels_breaks("fgco2"),
#         limits = c(quantile(.x$resid, .01), quantile(.x$resid, .99)),
#         oob = squish
#       ) +
#       theme(legend.title = element_markdown(), 
#             legend.position = "bottom") +
#       facet_grid(month ~ name,
#                  labeller = labeller(name = x_axis_labels)) +
#       guides(
#         fill = guide_colorbar(
#           barheight = unit(0.3, "cm"),
#           barwidth = unit(6, "cm"),
#           ticks = TRUE,
#           ticks.colour = "grey20",
#           frame.colour = "grey20",
#           label.position = "top",
#           direction = "horizontal"
#         )
#       ) +
#       theme(legend.title = element_markdown(),
#             legend.position = "top",
#             strip.text.x.top = element_markdown())
#   )


```

```{r flux_attribution_annual_maps, fig.asp=1.8}

pco2_product_map_annual_flux_attribution <-
  pco2_product_map_monthly_flux_attribution %>% 
  group_by(year, lat, lon, name) %>% 
  summarise(resid = mean(resid, na.rm = TRUE)) %>% 
  ungroup()

map +
  geom_tile(data = pco2_product_map_annual_flux_attribution %>%
              filter(year == {{year_anom}}), aes(lon, lat, fill = resid)) +
  scale_fill_gradientn(
    colours = cmocean("curl")(100),
    rescaler = ~ scales::rescale_mid(.x, mid = 0),
    name = labels_breaks("fgco2"),
    limits = c(
      quantile(pco2_product_map_annual_flux_attribution$resid, .01, na.rm = TRUE),
      quantile(pco2_product_map_annual_flux_attribution$resid, .99, na.rm = TRUE)
    ),
    oob = squish
  ) +
  theme(legend.title = element_markdown(), legend.position = "bottom") +
  facet_wrap(~ name,
             ncol = 2,
             labeller = labeller(name = x_axis_labels)) +
  guides(
    fill = guide_colorbar(
      barheight = unit(0.3, "cm"),
      barwidth = unit(6, "cm"),
      ticks = TRUE,
      ticks.colour = "grey20",
      frame.colour = "grey20",
      label.position = "top",
      direction = "horizontal"
    )
  ) +
  theme(
    legend.title = element_markdown(),
    legend.position = "top",
    strip.text.x.top = element_markdown()
    
  )
gc()

```


```{r write_map_monthly_anomaly, results = FALSE}

pco2_product_map_monthly_anomaly %>%
  filter(year == {{year_anom}}) %>%
  write_csv(
    paste0(
      "../data/",
      "{{product_name}}",
      "_",
      "{{year_anom}}",
      "_map_monthly_anomaly.csv"
    )
  )

pco2_product_map_annual_flux_attribution %>%
  filter(year == {{year_anom}}) %>%
  write_csv(
    paste0(
      "../data/",
      "{{product_name}}",
      "_",
      "{{year_anom}}",
      "_map_annual_flux_attribution.csv"
    )
  )

pco2_product_map_annual_fCO2_decomposition %>%
  filter(year == {{year_anom}}) %>%
  write_csv(
    paste0(
      "../data/",
      "{{product_name}}",
      "_",
      "{{year_anom}}",
      "_map_annual_fCO2_decomposition.csv"
    )
  )

pco2_product_map_monthly_flux_attribution %>%
  filter(year == {{year_anom}}) %>%
  write_csv(
    paste0(
      "../data/",
      "{{product_name}}",
      "_",
      "{{year_anom}}",
      "_map_monthly_flux_attribution.csv"
    )
  )

pco2_product_map_monthly_fCO2_decomposition %>%
  filter(year == {{year_anom}}) %>%
  write_csv(
    paste0(
      "../data/",
      "{{product_name}}",
      "_",
      "{{year_anom}}",
      "_map_monthly_fCO2_decomposition.csv"
    )
  )

rm(pco2_product_map_annual_flux_attribution,
   pco2_product_map_annual_fCO2_decomposition)

gc()


```



# Hovmoeller plots

The following Hovmoeller plots show the anomalies from the prediction of the linear/quadratic fits.

Hovmoeller plots are first presented as annual means, and than as monthly means. Note that the predictions for the monthly Hovmoeller plots are done individually for each month, such the mean seasonal anomaly from the annual mean is removed.

## {{year_anom}} annual anomalies

```{r {{year_anom}}_hovmoeller_annual_anomaly, fig.asp=0.7, results = FALSE}

pco2_product_hovmoeller_annual_anomaly <-
  pco2_product_hovmoeller_annual %>%
  anomaly_determination(lat) %>% 
  filter(!is.na(resid))

  
pco2_product_hovmoeller_annual_anomaly %>%
  # filter(name == "mld") %>%
  group_split(name) %>%
  head(1) %>%
  map(
    ~ ggplot(data = .x, aes(year, lat, fill = resid)) +
      geom_raster() +
      scale_fill_gradientn(
        colours = cmocean("curl")(100),
        rescaler = ~ scales::rescale_mid(.x, mid = 0),
        name = labels_breaks(.x %>% distinct(name)),
        limits = c(quantile(.x$resid, .01), quantile(.x$resid, .99)),
        oob = squish
      ) +
      coord_cartesian(expand = 0) +
      labs(title = "Annual mean anomalies", y = "Latitude") +
      guides(
        fill = guide_colorbar(
          barheight = unit(0.3, "cm"),
          barwidth = unit(6, "cm"),
          ticks = TRUE,
          ticks.colour = "grey20",
          frame.colour = "grey20",
          label.position = "top",
          direction = "horizontal"
        )
      ) +
      theme(legend.title = element_markdown(), legend.position = "top")
  )


```


## {{year_anom}} monthly anomalies

```{r {{year_anom}}_hovmoeller_monthly_anomalies, fig.asp=0.6, results = FALSE}

pco2_product_hovmoeller_monthly_anomaly <-
  pco2_product_hovmoeller_monthly %>%
  select(-c(decimal)) %>% 
  anomaly_determination(lat, month) %>% 
  filter(!is.na(resid))

  
pco2_product_hovmoeller_monthly_anomaly <-
  pco2_product_hovmoeller_monthly_anomaly %>%
  mutate(decimal = year + (month - 1) / 12)
  
pco2_product_hovmoeller_monthly_anomaly %>%
  group_split(name) %>%
  head(1) %>%
  map(
    ~ ggplot(data = .x,
             aes(decimal, lat, fill = resid)) +
      geom_raster() +
      scale_fill_gradientn(
        colours = cmocean("curl")(100),
        rescaler = ~ scales::rescale_mid(.x, mid = 0),
        name = labels_breaks(.x %>% distinct(name)),
        limits = c(quantile(.x$resid, .01), quantile(.x$resid, .99)),
        oob = squish
      ) +
      coord_cartesian(expand = 0) +
      labs(title = "Monthly mean anomalies",
           y = "Latitude") +
      guides(
        fill = guide_colorbar(
          barheight = unit(0.3, "cm"),
          barwidth = unit(6, "cm"),
          ticks = TRUE,
          ticks.colour = "grey20",
          frame.colour = "grey20",
          label.position = "top",
          direction = "horizontal"
        )
      ) +
      theme(legend.title = element_markdown(), legend.position = "top")
  )
  

```


### Three years prior {{year_anom}}

```{r {{year_anom}}_hovmoeller_monthly_anomalies_recent_years, fig.asp=0.6, results = FALSE}

pco2_product_hovmoeller_monthly_anomaly %>%
  filter(between(year, {{year_anom}}-2, {{year_anom}})) %>%
  group_split(name) %>%
  head(1) %>%
  map(
    ~ ggplot(data = .x,
             aes(decimal, lat, fill = resid)) +
      geom_raster() +
      scale_fill_gradientn(
        colours = cmocean("curl")(100),
        rescaler = ~ scales::rescale_mid(.x, mid = 0),
        name = labels_breaks(.x %>% distinct(name)),
        limits = c(quantile(.x$resid, .01), quantile(.x$resid, .99)),
        oob = squish
      ) +
      coord_cartesian(expand = 0) +
      labs(title = "Monthly mean anomalies",
           y = "Latitude") +
      guides(
        fill = guide_colorbar(
          barheight = unit(0.3, "cm"),
          barwidth = unit(6, "cm"),
          ticks = TRUE,
          ticks.colour = "grey20",
          frame.colour = "grey20",
          label.position = "top",
          direction = "horizontal"
        )
      ) +
      theme(legend.title = element_markdown(), legend.position = "top")
  )



```

```{r write_hovmoeller_monthly_anomalies, results = FALSE}

pco2_product_hovmoeller_monthly_anomaly %>%
  write_csv(
    paste0(
      "../data/",
      "{{product_name}}",
      "_",
      "{{year_anom}}",
      "_anomaly_hovmoeller_monthly.csv"
    )
  )

rm(
  pco2_product_hovmoeller_annual,
  pco2_product_hovmoeller_monthly,
  pco2_product_hovmoeller_annual_anomaly,
  pco2_product_hovmoeller_monthly_anomaly
)

gc()

```


# Regional means and integrals

The following plots show regionally averaged (or integrated) values of each variable as provided through the pCO2 product, as well as the anomalies from the prediction of a linear/quadratic fit.

Anomalies are first presented relative to the predicted annual mean of each year, hence preserving the seasonality. Furthermore, anomalies are presented relative to the predicted monthly mean values, such that the mean seasonality is removed.


## {{year_anom}} absolute values

### Global

```{r figure_height}

fig.height <- pco2_product_biome_monthly %>% 
  distinct(name) %>% 
  nrow()

fig.height <- (fig.height + 2) * 0.1

```



```{r {{year_anom}}_absolute_means_and_integrals_global, fig.asp=fig.height}

pco2_product_biome_monthly %>%
  filter(biome %in% "Global") %>%
  ggplot(aes(month, value, group = as.factor(year))) +
  geom_path(data = . %>% filter(!between(year, {{year_anom}}-1, {{year_anom}})),
            aes(col = year)) +
  scale_color_grayC() +
  new_scale_color() +
  geom_path(data = . %>% filter(between(year, {{year_anom}}-1, {{year_anom}})),
            aes(col = as.factor(year)),
            linewidth = 1) +
  scale_color_manual(values = c("orange", "red"),
                     guide = guide_legend(reverse = TRUE,
                                          order = 1)) +
  scale_x_continuous(breaks = seq(1, 12, 3), expand = c(0, 0)) +
  labs(title = "Absolute values | Global") +
  facet_wrap(name ~ .,
             scales = "free_y",
             labeller = labeller(name = x_axis_labels),
             strip.position = "left",
             ncol = 2) +
  theme(
    strip.text.y.left = element_markdown(),
    strip.placement = "outside",
    strip.background.y = element_blank(),
    legend.title = element_blank(),
    axis.title.y = element_blank()
  )


```

### Key biomes

```{r {{year_anom}}_absolute_means_and_integrals_overview_biomes, fig.asp=fig.height}

pco2_product_biome_monthly %>%
  filter(biome %in% key_biomes) %>%
  ggplot(aes(month, value, group = as.factor(year))) +
  geom_path(data = . %>% filter(!between(year, {{year_anom}}-1, {{year_anom}})),
            aes(col = year)) +
  scale_color_grayC() +
  new_scale_color() +
  geom_path(data = . %>% filter(between(year, {{year_anom}}-1, {{year_anom}})),
            aes(col = as.factor(year)),
            linewidth = 1) +
  scale_color_manual(values = c("orange", "red"),
                     guide = guide_legend(reverse = TRUE,
                                          order = 1)) +
  scale_x_continuous(breaks = seq(1, 12, 3), expand = c(0, 0)) +
  labs(title = "Absolute values | Selected biomes") +
  facet_grid(name ~ biome,
             scales = "free_y",
             labeller = labeller(name = x_axis_labels),
             switch = "y") +
  theme(
    strip.text.y.left = element_markdown(),
    strip.placement = "outside",
    strip.background.y = element_blank(),
    legend.title = element_blank(),
    axis.title.y = element_blank()
  )


```


```{r {{year_anom}}_absolute_means_and_integrals_key_biomes, fig.asp=fig.height, results = FALSE}

pco2_product_biome_monthly %>%
  filter(biome %in% key_biomes) %>%
  group_split(biome) %>%
  head(1) %>%
  map(
    ~ ggplot(data = .x,
             aes(month, value, group = as.factor(year))) +
      geom_path(data = . %>% filter(!between(year, {{year_anom}}-1, {{year_anom}})),
                aes(col = year)) +
      scale_color_grayC() +
      new_scale_color() +
      geom_path(
        data = . %>% filter(between(year, {{year_anom}}-1, {{year_anom}})),
        aes(col = as.factor(year)),
        linewidth = 1
      ) +
      scale_color_manual(
        values = c("orange", "red"),
        guide = guide_legend(reverse = TRUE,
                             order = 1)
      ) +
      scale_x_continuous(breaks = seq(1, 12, 3), expand = c(0, 0)) +
  labs(title = paste("Absolute values |", .x$biome)) +
  facet_wrap(name ~ .,
             scales = "free_y",
             labeller = labeller(name = x_axis_labels),
             strip.position = "left",
             ncol = 2) +
  theme(
    strip.text.y.left = element_markdown(),
    strip.placement = "outside",
    strip.background.y = element_blank(),
    legend.title = element_blank(),
    axis.title.y = element_blank()
  )
  )

```


## {{year_anom}} anomalies

### Annual mean trends

```{r {{year_anom}}_annual_mean_trends, fig.asp=fig.height}

pco2_product_biome_annual %>% 
  filter(biome %in% "Global") %>%
  ggplot(aes(year, value)) +
  geom_path() +
  geom_point(data = . %>% filter(year != {{year_anom}})) +
  geom_point(data = . %>% filter(year == {{year_anom}}),
             shape = 1) +
  geom_smooth(data = . %>% filter(year != {{year_anom}},
                                  !(name %in% name_quadratic_fit)),
              method = "lm",
              fullrange = TRUE,
              aes(col = "linear"),
              se = FALSE) +
  geom_smooth(data = . %>% filter(year != {{year_anom}},
                                  name %in% name_quadratic_fit),
              method = "lm",
              fullrange = TRUE,
              formula = y ~ x + I(x^2),
              aes(col = "quadratic"),
              se = FALSE) +
  scale_color_brewer(
    palette = "Set1",
    name = paste("Regression fit\nexcl.", {{year_anom}})) +
  scale_x_continuous(breaks = seq(1980, 2020, 20)) +
  labs(title = "Annual mean trends | Global") +
  facet_wrap(name ~ .,
             scales = "free_y",
             labeller = labeller(name = x_axis_labels),
             strip.position = "left",
             ncol = 2) +
  theme(
    strip.text.y.left = element_markdown(),
    strip.placement = "outside",
    strip.background.y = element_blank(),
    axis.title = element_blank()
  )
  

pco2_product_biome_annual %>% 
  filter(biome %in% key_biomes) %>%
  ggplot(aes(year, value)) +
  geom_path() +
  geom_point(data = . %>% filter(year != {{year_anom}}),
             size = 0.2) +
  geom_point(data = . %>% filter(year == {{year_anom}}),
             shape = 1) +
  geom_smooth(data = . %>% filter(year != {{year_anom}},
                                  !(name %in% name_quadratic_fit)),
              method = "lm",
              fullrange = TRUE,
              aes(col = "linear"),
              se = FALSE) +
  geom_smooth(data = . %>% filter(year != {{year_anom}},
                                  name %in% name_quadratic_fit),
              method = "lm",
              fullrange = TRUE,
              formula = y ~ x + I(x^2),
              aes(col = "quadratic"),
              se = FALSE) +
  scale_color_brewer(
    palette = "Set1",
    name = paste("Regression fit\nexcl.", {{year_anom}})) +
  scale_x_continuous(breaks = seq(1980, 2020, 20)) +
  labs(title = "Annual mean trends | Selected biomes") +
  facet_grid(name ~ biome,
             scales = "free_y",
             labeller = labeller(name = x_axis_labels),
             switch = "y") +
  theme(
    strip.text.y.left = element_markdown(),
    strip.placement = "outside",
    strip.background.y = element_blank(),
    axis.title = element_blank()
  )

pco2_product_biome_annual_anomaly <-
  pco2_product_biome_annual %>%
  anomaly_determination(biome)
  

pco2_product_biome_annual_anomaly %>%
  filter(biome %in% "Global") %>%
  ggplot(aes(year, resid)) +
  geom_hline(yintercept = 0) +
  geom_path() +
  geom_point(data = . %>% filter(!between(year, {{year_anom}}-1, {{year_anom}})),
             aes(fill = year),
             shape = 21) +
  scale_fill_grayC() +
  new_scale_fill() +
  geom_point(data = . %>% filter(between(year, {{year_anom}}-1, {{year_anom}})),
             aes(fill = as.factor(year)),
             shape = 21, size = 2)  +
  scale_fill_manual(values = c("orange", "red"),
                     guide = guide_legend(reverse = TRUE,
                                          order = 1)) +
  scale_x_continuous(breaks = seq(1980, 2020, 20)) +
  labs(title = "Residual from fit to annual means (actual - predicted) | Global") +
  facet_wrap(name ~ .,
             scales = "free_y",
             labeller = labeller(name = x_axis_labels),
             strip.position = "left",
             ncol = 2) +
  theme(
    strip.text.y.left = element_markdown(),
    strip.placement = "outside",
    strip.background.y = element_blank(),
    axis.title = element_blank(),
    legend.title = element_blank()
  )
  

pco2_product_biome_annual_anomaly %>%
  filter(biome %in% key_biomes) %>%
  ggplot(aes(year, resid)) +
  geom_hline(yintercept = 0) +
  geom_path() +
  geom_point(data = . %>% filter(!between(year, {{year_anom}}-1, {{year_anom}})),
             aes(fill = year),
             shape = 21) +
  scale_fill_grayC() +
  new_scale_fill() +
  geom_point(data = . %>% filter(between(year, {{year_anom}}-1, {{year_anom}})),
             aes(fill = as.factor(year)),
             shape = 21, size = 2)  +
  scale_fill_manual(values = c("orange", "red"),
                     guide = guide_legend(reverse = TRUE,
                                          order = 1)) +
  scale_x_continuous(breaks = seq(1980, 2020, 20)) +
  labs(title = "Residual from fit to annual means (actual - predicted) | Selected biomes") +
  facet_grid(name ~ biome,
             scales = "free_y",
             labeller = labeller(name = x_axis_labels),
             switch = "y") +
  theme(
    strip.text.y.left = element_markdown(),
    strip.placement = "outside",
    strip.background.y = element_blank(),
    axis.title = element_blank(),
    legend.title = element_blank()
  )


```


```{r write_biome_annual_anomaly}

pco2_product_biome_annual_anomaly %>%
  write_csv(
    paste0(
      "../data/",
      "{{product_name}}",
      "_",
      "{{year_anom}}",
      "_biome_annual_anomaly.csv"
    )
  )

```



```{r {{year_anom}}_anomalies_from_annual_mean_trends_overview, fig.asp=fig.height}

pco2_product_biome_annual_detrended <-
  full_join(pco2_product_biome_monthly,
            pco2_product_biome_annual_anomaly %>% select(-c(value, resid))) %>%
  mutate(resid = value - fit)

pco2_product_biome_annual_detrended %>% 
  filter(biome %in% "Global") %>%
  ggplot(aes(month, resid, group = as.factor(year))) +
  geom_path(data = . %>% filter(!between(year, {{year_anom}}-1, {{year_anom}})),
            aes(col = year)) +
  scale_color_grayC() +
  new_scale_color() +
  geom_path(data = . %>% filter(between(year, {{year_anom}}-1, {{year_anom}})),
            aes(col = as.factor(year)),
            linewidth = 1) +
  scale_color_manual(values = c("orange", "red"),
                     guide = guide_legend(reverse = TRUE,
                                          order = 1)) +
  scale_x_continuous(breaks = seq(1, 12, 3), expand = c(0, 0)) +
  labs(title = "Anomalies from predicted annual mean | Global") +
  facet_wrap(name ~ .,
             scales = "free_y",
             labeller = labeller(name = x_axis_labels),
             strip.position = "left",
             ncol = 2) +
  theme(
    strip.text.y.left = element_markdown(),
    strip.placement = "outside",
    strip.background.y = element_blank(),
    axis.title.y = element_blank(),
    legend.title = element_blank()
  )

pco2_product_biome_annual_detrended %>% 
  filter(biome %in% key_biomes) %>%
  ggplot(aes(month, resid, group = as.factor(year))) +
  geom_path(data = . %>% filter(!between(year, {{year_anom}}-1, {{year_anom}})),
            aes(col = year)) +
  scale_color_grayC() +
  new_scale_color() +
  geom_path(data = . %>% filter(between(year, {{year_anom}}-1, {{year_anom}})),
            aes(col = as.factor(year)),
            linewidth = 1) +
  scale_color_manual(values = c("orange", "red"),
                     guide = guide_legend(reverse = TRUE,
                                          order = 1)) +
  scale_x_continuous(breaks = seq(1, 12, 3), expand = c(0, 0)) +
  labs(title = "Anomalies from predicted annual mean | Selected biomes") +
  facet_grid(name ~ biome,
             scales = "free_y",
             labeller = labeller(name = x_axis_labels),
             switch = "y") +
  theme(
    strip.text.y.left = element_markdown(),
    strip.placement = "outside",
    strip.background.y = element_blank(),
    axis.title.y = element_blank(),
    legend.title = element_blank(),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)
  )

```


```{r {{year_anom}}_anomalies_from_annual_mean_trends_key_biomes, fig.asp=fig.height, results = FALSE}

pco2_product_biome_annual_detrended %>%
  filter(biome %in% key_biomes) %>%
  group_split(biome) %>%
  head(1) %>%
  map(
    ~ ggplot(data = .x,
             aes(month, resid, group = as.factor(year))) +
      geom_path(data = . %>% filter(!between(year, {{year_anom}}-1, {{year_anom}})),
                aes(col = year)) +
      scale_color_grayC() +
      new_scale_color() +
      geom_path(
        data = . %>% filter(between(year, {{year_anom}}-1, {{year_anom}})),
        aes(col = as.factor(year)),
        linewidth = 1
      ) +
      scale_color_manual(
        values = c("orange", "red"),
        guide = guide_legend(reverse = TRUE,
                             order = 1)
      ) +
      scale_x_continuous(breaks = seq(1, 12, 3), expand = c(0, 0)) +
      labs(title = paste("Anomalies from predicted annual mean |", .x$biome)) +
      facet_wrap(
        name ~ .,
        scales = "free_y",
        labeller = labeller(name = x_axis_labels),
        strip.position = "left",
        ncol = 2
      ) +
      theme(
        strip.text.y.left = element_markdown(),
        strip.placement = "outside",
        strip.background.y = element_blank(),
        axis.title = element_blank(),
        legend.title = element_blank()
      )
  )


```

```{r write_biome_annual_detrended}

pco2_product_biome_annual_detrended %>%
  write_csv(
    paste0(
      "../data/",
      "{{product_name}}",
      "_",
      "{{year_anom}}",
      "_biome_annual_detrended.csv"
    )
  )

```



### Monthly mean trends

```{r {{year_anom}}_monthly_mean_trends, fig.asp=fig.height}

# pco2_product_biome_monthly %>% 
#   filter(biome %in% "Global") %>%
#   mutate(month = as.factor(month)) %>% 
#   ggplot(aes(year, value, col = month)) +
#   # geom_point() +
#   geom_smooth(data = . %>% filter(year != {{year_anom}},
#                                   !(name %in% name_quadratic_fit)),
#               method = "lm",
#               se = FALSE,
#               fullrange = TRUE) +
#   geom_smooth(
#     data = . %>% filter(year != {{year_anom}},
#                         name %in% name_quadratic_fit),
#     method = "lm",
#     fullrange = TRUE,
#     formula = y ~ x + I(x ^ 2),
#     se = FALSE
#   ) +
#   scale_color_scico_d(palette = "romaO", 
#     name = paste("Regression fit\nexcl.", {{year_anom}})) +
#   scale_x_continuous(breaks = seq(1980, 2020, 20)) +
#   labs(title = "Monthly mean trends | Global") +
#   facet_wrap(name ~ .,
#              scales = "free_y",
#              labeller = labeller(name = x_axis_labels),
#              strip.position = "left",
#              ncol = 2) +
#   theme(
#     strip.text.y.left = element_markdown(),
#     strip.placement = "outside",
#     strip.background.y = element_blank(),
#     axis.title = element_blank()
#   )
# 
# pco2_product_biome_monthly %>% 
#   filter(biome %in% key_biomes) %>%
#   mutate(month = as.factor(month)) %>% 
#   ggplot(aes(year, value, col = month)) +
#   # geom_point() +
#   geom_smooth(data = . %>% filter(year != {{year_anom}},
#                                   !(name %in% name_quadratic_fit)),
#               method = "lm",
#               fullrange = TRUE,
#               se = FALSE) +
#   geom_smooth(
#     data = . %>% filter(year != {{year_anom}},
#                         name %in% name_quadratic_fit),
#     method = "lm",
#     fullrange = TRUE,
#     formula = y ~ x + I(x ^ 2),
#     se = FALSE
#   ) +
#   scale_color_scico_d(palette = "romaO",
#     name = paste("Regression fit\nexcl.", {{year_anom}})) +
#   scale_x_continuous(breaks = seq(1980, 2020, 20)) +
#   labs(title = "Monthly mean trends | Selected biomes") +
#   facet_grid(name ~ biome,
#              scales = "free_y",
#              labeller = labeller(name = x_axis_labels),
#              switch = "y") +
#   theme(
#     strip.text.y.left = element_markdown(),
#     strip.placement = "outside",
#     strip.background.y = element_blank(),
#     axis.title = element_blank()
#   )

pco2_product_biome_monthly_anomaly <-
  pco2_product_biome_monthly %>%
  anomaly_determination(biome, month)


# pco2_product_biome_monthly_anomaly %>%
#   filter(biome %in% "Global") %>%
#   group_split(month) %>%
#   head(1) %>%
#   map(
#     ~ ggplot(data = .x,
#              aes(year, resid)) +
#       geom_hline(yintercept = 0) +
#       geom_path() +
#       geom_point(
#         data = . %>% filter(!between(year, {{year_anom}}-1, {{year_anom}})),
#         aes(fill = year),
#         shape = 21
#       ) +
#       scale_fill_grayC() +
#       new_scale_fill() +
#       geom_point(
#         data = . %>% filter(between(year, {{year_anom}}-1, {{year_anom}})),
#         aes(fill = as.factor(year)),
#         shape = 21,
#         size = 2
#       )  +
#       scale_fill_manual(
#         values = c("orange", "red"),
#         guide = guide_legend(reverse = TRUE,
#                              order = 1)
#       ) +
#       scale_x_continuous(breaks = seq(1980, 2020, 20)) +
#       labs(title = "Residual from fit to monthly means (actual - predicted) | Global",
#            subtitle = paste("Month:", .x$month)) +
#       facet_wrap(
#         name ~ .,
#         scales = "free_y",
#         labeller = labeller(name = x_axis_labels),
#         strip.position = "left",
#         ncol = 2
#       ) +
#       theme(
#         strip.text.y.left = element_markdown(),
#         strip.placement = "outside",
#         strip.background.y = element_blank(),
#         axis.title = element_blank(),
#         legend.title = element_blank()
#       )
#   )
# 
# pco2_product_biome_monthly_anomaly %>%
#   filter(biome %in% key_biomes) %>%
#   group_split(month) %>%
#   head(1) %>%
#   map(
#     ~ ggplot(data = .x,
#              aes(year, resid)) +
#       geom_hline(yintercept = 0) +
#       geom_path() +
#       geom_point(
#         data = . %>% filter(!between(year, {{year_anom}}-1, {{year_anom}})),
#         aes(fill = year),
#         shape = 21
#       ) +
#       scale_fill_grayC() +
#       new_scale_fill() +
#       geom_point(
#         data = . %>% filter(between(year, {{year_anom}}-1, {{year_anom}})),
#         aes(fill = as.factor(year)),
#         shape = 21,
#         size = 2
#       )  +
#       scale_fill_manual(
#         values = c("orange", "red"),
#         guide = guide_legend(reverse = TRUE,
#                              order = 1)
#       ) +
#       scale_x_continuous(breaks = seq(1980, 2020, 20)) +
#       labs(title = "Residual from fit to monthly means (actual - predicted) | Selected biomes",
#            subtitle = paste("Month:", .x$month)) +
#       facet_grid(
#         name ~ biome,
#         scales = "free_y",
#         labeller = labeller(name = x_axis_labels),
#         switch = "y"
#       ) +
#       theme(
#         strip.text.y.left = element_markdown(),
#         strip.placement = "outside",
#         strip.background.y = element_blank(),
#         axis.title = element_blank(),
#         legend.title = element_blank()
#       )
#   )


```


```{r write_biome_monthly_anomaly}

pco2_product_biome_monthly_anomaly %>%
  write_csv(
    paste0(
      "../data/",
      "{{product_name}}",
      "_",
      "{{year_anom}}",
      "_biome_monthly_anomaly.csv"
    )
  )

```



### Global

```{r {{year_anom}}_anomalies_from_monthly_mean_trends_global, fig.asp=fig.height}


pco2_product_biome_monthly_detrended <-
  full_join(pco2_product_biome_monthly,
            pco2_product_biome_monthly_anomaly %>% select(-c(value, resid))) %>%
  mutate(resid = value - fit)

pco2_product_biome_monthly_detrended %>% 
  filter(biome %in% "Global") %>%
  ggplot(aes(month, resid, group = as.factor(year))) +
  geom_path(data = . %>% filter(!between(year, {{year_anom}}-1, {{year_anom}})),
            aes(col = year)) +
  scale_color_grayC() +
  new_scale_color() +
  geom_path(data = . %>% filter(between(year, {{year_anom}}-1, {{year_anom}})),
            aes(col = as.factor(year)),
            linewidth = 1) +
  scale_color_manual(values = c("orange", "red"),
                     guide = guide_legend(reverse = TRUE,
                                          order = 1)) +
  scale_x_continuous(breaks = seq(1, 12, 3), expand = c(0, 0)) +
  labs(title = "Anomalies from predicted monthly mean | Global") +
  facet_wrap(
    name ~ .,
    scales = "free_y",
    labeller = labeller(name = x_axis_labels),
    strip.position = "left",
    ncol = 2
  ) +
  theme(
    strip.text.y.left = element_markdown(),
    strip.placement = "outside",
    strip.background.y = element_blank(),
    axis.title.y = element_blank(),
    legend.title = element_blank()
  )

pco2_product_biome_monthly_detrended %>% 
  filter(biome %in% key_biomes) %>%
  ggplot(aes(month, resid, group = as.factor(year))) +
  geom_path(data = . %>% filter(!between(year, {{year_anom}}-1, {{year_anom}})),
            aes(col = year)) +
  scale_color_grayC() +
  new_scale_color() +
  geom_path(data = . %>% filter(between(year, {{year_anom}}-1, {{year_anom}})),
            aes(col = as.factor(year)),
            linewidth = 1) +
  scale_color_manual(values = c("orange", "red"),
                     guide = guide_legend(reverse = TRUE,
                                          order = 1)) +
  scale_x_continuous(breaks = seq(1, 12, 3), expand = c(0, 0)) +
  labs(title = "Anomalies from predicted monthly mean | Selected biomes") +
  facet_grid(
    name ~ biome,
    scales = "free_y",
    labeller = labeller(name = x_axis_labels),
    switch = "y"
  ) +
  theme(
    strip.text.y.left = element_markdown(),
    strip.placement = "outside",
    strip.background.y = element_blank(),
    axis.title.y = element_blank(),
    legend.title = element_blank()
  )

```

### Key biomes

```{r {{year_anom}}_anomalies_from_monthly_mean_trends__key_biomes_overview, fig.asp=fig.height, results = FALSE}

pco2_product_biome_monthly_detrended %>% 
  filter(biome %in% key_biomes) %>%
  ggplot(aes(month, resid, group = as.factor(year))) +
  geom_path(data = . %>% filter(!between(year, {{year_anom}}-1, {{year_anom}})),
            aes(col = year)) +
  scale_color_grayC() +
  new_scale_color() +
  geom_path(data = . %>% filter(between(year, {{year_anom}}-1, {{year_anom}})),
            aes(col = as.factor(year)),
            linewidth = 1) +
  scale_color_manual(values = c("orange", "red"),
                     guide = guide_legend(reverse = TRUE,
                                          order = 1)) +
  scale_x_continuous(breaks = seq(1, 12, 3), expand = c(0, 0)) +
  labs(title = "Anomalies from predicted monthly mean | Selected biomes") +
  facet_grid(
    name ~ biome,
    scales = "free_y",
    labeller = labeller(name = x_axis_labels),
    switch = "y"
  ) +
  theme(
    strip.text.y.left = element_markdown(),
    strip.placement = "outside",
    strip.background.y = element_blank(),
    axis.title.y = element_blank(),
    legend.title = element_blank()
  )

```


```{r {{year_anom}}_anomalies_from_monthly_mean_trends_key_biomes, fig.asp=fig.height, results = FALSE}

pco2_product_biome_monthly_detrended %>%
  filter(biome %in% key_biomes) %>%
  group_split(biome) %>%
  head(1) %>%
  map(
    ~ ggplot(data = .x,
             aes(month, resid, group = as.factor(year))) +
      geom_path(data = . %>% filter(!between(year, {{year_anom}}-1, {{year_anom}})),
                aes(col = year)) +
      scale_color_grayC() +
      new_scale_color() +
      geom_path(
        data = . %>% filter(between(year, {{year_anom}}-1, {{year_anom}})),
        aes(col = as.factor(year)),
        linewidth = 1
      ) +
      scale_color_manual(
        values = c("orange", "red"),
        guide = guide_legend(reverse = TRUE,
                             order = 1)
      ) +
      scale_x_continuous(breaks = seq(1, 12, 3), expand = c(0, 0)) +
      labs(title = paste("Anomalies from predicted monthly mean |", .x$biome)) +
      facet_wrap(
        name ~ .,
        scales = "free_y",
        labeller = labeller(name = x_axis_labels),
        strip.position = "left",
        ncol = 2
      ) +
      theme(
        strip.text.y.left = element_markdown(),
        strip.placement = "outside",
        strip.background.y = element_blank(),
        axis.title.y = element_blank(),
        legend.title = element_blank()
      )
  )


```

```{r write_biome_monthly_detrended}

pco2_product_biome_monthly_detrended %>%
  write_csv(
    paste0(
      "../data/",
      "{{product_name}}",
      "_",
      "{{year_anom}}",
      "_biome_monthly_detrended.csv"
    )
  )

```




# {{year_anom}} anomaly correlation


The following plots aim to unravel the correlation between regionally integrated monthly flux anomalies and the corresponding anomalies of the means/integrals of each other variable.

Anomalies are first presented are first presented in absolute units. Due to the different flux magnitudes, we need to plot integrated fluxes separately for each region. Secondly, we normalize the monthly anomalies to the spread (expressed as standard deviation) of the residuals from the fit.

## Annual anomalies

### Absolute

```{r {{year_anom}}_anomalies_correlation_absolute_global_annual, fig.asp=fig.height}

pco2_product_biome_annual_anomaly %>%
  filter(biome == "Global") %>%
  select(-c(value, fit)) %>% 
  pivot_wider(values_from = resid) %>% 
  pivot_longer(-c(year, biome, fgco2_int))  %>%
  ggplot(aes(value, fgco2_int)) +
  geom_hline(yintercept = 0) +
  geom_point(data = . %>% filter(!between(year, {{year_anom}}-1, {{year_anom}})),
             aes(fill = year),
             shape = 21) +
  geom_smooth(
    data = . %>% filter(!between(year, {{year_anom}}-1, {{year_anom}})),
    method = "lm",
    se = FALSE,
    fullrange = TRUE,
    aes(col = paste("Regression fit\nexcl.", {{year_anom}}))
  ) +
  scale_color_grey() +
  scale_fill_grayC()+
  new_scale_fill() +
  geom_point(data = . %>% filter(between(year, {{year_anom}}-1, {{year_anom}})),
             aes(fill = as.factor(year)),
             shape = 21, size = 2)  +
  scale_fill_manual(values = c("orange", "red"),
                     guide = guide_legend(reverse = TRUE,
                                          order = 1)) +
  labs(title = "Globally integrated fluxes",
       y = labels_breaks("fgco2_int")$i_legend_title) +
  facet_wrap(
    ~ name,
    scales = "free_x",
    labeller = labeller(name = x_axis_labels),
    strip.position = "bottom",
    ncol = 2
  ) +
  theme(
    strip.text.x.bottom = element_markdown(),
    strip.placement = "outside",
    strip.background.x = element_blank(),
    axis.title.y = element_markdown(),
    axis.title.x = element_blank(),
    legend.title = element_blank()
  )


```

## Monthly anomalies

### Absolute

```{r {{year_anom}}_anomalies_correlation_absolute_global_monthly, fig.asp=fig.height}


pco2_product_biome_monthly_detrended_anomaly <-
  pco2_product_biome_monthly_detrended %>%
  select(year, month, biome, name, resid) %>%
  pivot_wider(names_from = name,
              values_from = resid)


pco2_product_biome_monthly_detrended_anomaly %>%
  filter(biome == "Global") %>%
  pivot_longer(-c(year, month, biome, fgco2_int))  %>%
  ggplot(aes(value, fgco2_int)) +
  geom_hline(yintercept = 0) +
  geom_point(data = . %>% filter(year != {{year_anom}}),
             aes(col = paste(min(year), max(year), sep = "-")),
             alpha = 0.2) +
  geom_smooth(
    data = . %>% filter(year != {{year_anom}}),
    aes(col = paste(min(year), max(year), sep = "-")),
    method = "lm",
    se = FALSE,
    fullrange = TRUE
  )  +
  scale_color_grey(name = "") +
  new_scale_color() +
  geom_path(data = . %>% filter(year == {{year_anom}}),
            aes(col = as.factor(month), group = 1))  +
  geom_point(data = . %>% filter(year == {{year_anom}}),
             aes(fill =  as.factor(month)),
             shape = 21,
             size = 3)  +
  scale_color_scico_d(palette = "buda",
                     guide = guide_legend(reverse = TRUE,
                                          order = 1),
                     name = paste("Month\nof", {{year_anom}})) +
  scale_fill_scico_d(palette = "buda",
                     guide = guide_legend(reverse = TRUE,
                                          order = 1),
                     name = paste("Month\nof", {{year_anom}})) +
  labs(title = "Globally integrated fluxes",
       y = labels_breaks("fgco2_int")$i_legend_title) +
  facet_wrap(
    ~ name,
    scales = "free_x",
    labeller = labeller(name = x_axis_labels),
    strip.position = "bottom",
    ncol = 2
  ) +
  theme(
    strip.text.x.bottom = element_markdown(),
    strip.placement = "outside",
    strip.background.x = element_blank(),
    axis.title.y = element_markdown(),
    axis.title.x = element_blank()
  )


```

```{r {{year_anom}}_anomalies_correlation_absolute_biome_monthly, fig.asp=1.8, results = FALSE}

pco2_product_biome_monthly_detrended_anomaly %>%
  filter(!(biome %in% c(key_biomes, "Global"))) %>%
  pivot_longer(-c(year, month, biome, fgco2_int))  %>%
  filter(name %in% c("temperature", "chl", "dfco2", "kw_sol")) %>% 
  group_split(name) %>%
  head(1) %>%
  map(
    ~ ggplot(data = .x,
             aes(value, fgco2_int)) +
      geom_hline(yintercept = 0) +
      geom_point(
        data = . %>% filter(year != {{year_anom}}),
        aes(col = paste(min(year), max(year), sep = "-")),
        alpha = 0.2
      ) +
      geom_smooth(
        data = . %>% filter(year != {{year_anom}}),
        aes(col = paste(min(year), max(year), sep = "-")),
        method = "lm",
        se = FALSE,
        fullrange = TRUE
      )  +
      scale_color_grey(name = "") +
      new_scale_color() +
      geom_path(data = . %>% filter(year == {{year_anom}}),
                aes(col = as.factor(month), group = 1))  +
      geom_point(
        data = . %>% filter(year == {{year_anom}}),
        aes(fill = as.factor(month)),
        shape = 21,
        size = 3
      )  +
      scale_color_scico_d(
        palette = "buda",
        guide = guide_legend(reverse = TRUE,
                             order = 1),
        name = paste("Month\nof", {{year_anom}})
      ) +
      scale_fill_scico_d(
        palette = "buda",
        guide = guide_legend(reverse = TRUE,
                             order = 1),
        name = paste("Month\nof", {{year_anom}})
      ) +
      facet_wrap( ~ biome, ncol = 3, scales = "free") +
      labs(
        title = "Biome integrated fluxes",
        y = labels_breaks("fgco2_int")$i_legend_title,
        x = labels_breaks(.x %>% distinct(name))$i_legend_title
      ) +
      theme(axis.title.x = element_markdown(),
            axis.title.y = element_markdown())
  )


```

### Relative to spread

```{r {{year_anom}}_anomalies_correlation_realtive_to_spread_monthly, fig.asp=2, results = FALSE}


pco2_product_biome_monthly_detrended_anomaly_spread <-
  pco2_product_biome_monthly_detrended_anomaly %>%
  pivot_longer(-c(month, biome, year)) %>%
  filter(year != {{year_anom}}) %>%
  group_by(month, biome, name) %>%
  summarise(spread = sd(value, na.rm = TRUE)) %>%
  ungroup()



pco2_product_biome_monthly_detrended_anomaly_relative <-
  full_join(
    pco2_product_biome_monthly_detrended_anomaly_spread,
    pco2_product_biome_monthly_detrended_anomaly %>%
      pivot_longer(-c(month, biome, year))
  )

pco2_product_biome_monthly_detrended_anomaly_relative <-
  pco2_product_biome_monthly_detrended_anomaly_relative %>%
  mutate(value = value / spread) %>%
  select(-spread) %>%
  pivot_wider() %>%
  pivot_longer(-c(month, biome, year, fgco2_int))



pco2_product_biome_monthly_detrended_anomaly_relative %>%
  filter(name %in% c("temperature", "chl", "dfco2", "kw_sol")) %>% 
  group_split(name) %>%
  head(1) %>%
  map(
    ~ ggplot(data = .x,
             aes(value, fgco2_int)) +
      geom_vline(xintercept = 0) +
      geom_hline(yintercept = 0) +
      geom_point(
        data = . %>% filter(year != {{year_anom}}),
        aes(col = paste(min(year), max(year), sep = "-")),
        alpha = 0.2
      ) +
      geom_smooth(
        data = . %>% filter(year != {{year_anom}}),
        aes(col = paste(min(year), max(year), sep = "-")),
        method = "lm",
        se = FALSE,
        fullrange = TRUE
      )  +
      scale_color_grey(name = "") +
      new_scale_color() +
      geom_path(data = . %>% filter(year == {{year_anom}}),
                aes(col = as.factor(month), group = 1))  +
      geom_point(
        data = . %>% filter(year == {{year_anom}}),
        aes(fill = as.factor(month)),
        shape = 21,
        size = 3
      )  +
      scale_color_scico_d(
        palette = "buda",
        guide = guide_legend(reverse = TRUE,
                             order = 1),
        name = paste("Month\nof", {{year_anom}})
      ) +
      scale_fill_scico_d(
        palette = "buda",
        guide = guide_legend(reverse = TRUE,
                             order = 1),
        name = paste("Month\nof", {{year_anom}})
      ) +
      facet_wrap( ~ biome, ncol = 3) +
      coord_fixed() +
      labs(
        title = "Biome integrated fluxes normalized to spread",
        y = str_split_i(labels_breaks("fgco2_int")$i_legend_title, "<br>", i = 1),
        x = str_split_i(labels_breaks(.x %>% distinct(name))$i_legend_title, "<br>", i = 1)
      ) +
      theme(axis.title.x = element_markdown(),
            axis.title.y = element_markdown())
  )


```

## fCO2 decomposition

```{r pCO2_decomposition}

biome_mask <-
  bind_rows(
    biome_mask,
    biome_mask %>% mutate(biome = "Global")
  )

pco2_product_biome_monthly_fCO2_decomposition <-
  full_join(pco2_product_map_monthly_fCO2_decomposition,
            biome_mask,
            relationship = "many-to-many") %>% 
  group_by(year, month, biome, name) %>% 
  summarise(resid = mean(resid, na.rm = TRUE)) %>% 
  ungroup() %>% 
  drop_na()


pco2_product_biome_annual_fCO2_decomposition <-
  pco2_product_biome_monthly_fCO2_decomposition %>%
  group_by(year, biome, name) %>%
  summarise(resid = mean(resid)) %>%
  ungroup()


pco2_product_biome_monthly_fCO2_decomposition %>%
  filter(biome %in% c("Global", key_biomes)) %>%
  p_season(title  = paste("Anomalies from predicted monthly mean"))


```

## Flux attribution

### Seasonal

```{r flux_attribution, fig.asp=1}

pco2_product_biome_monthly_flux_attribution <-
  full_join(pco2_product_map_monthly_flux_attribution,
            biome_mask,
            relationship = "many-to-many") %>% 
  group_by(year, month, biome, name) %>% 
  summarise(resid = mean(resid, na.rm = TRUE)) %>% 
  ungroup() %>% 
  drop_na()

pco2_product_biome_monthly_flux_attribution_total <-
  full_join(pco2_product_map_monthly_anomaly %>% 
              filter(name == "fgco2") %>% 
              mutate(name = "resid_fgco2"),
            biome_mask,
            relationship = "many-to-many") %>% 
  group_by(year, month, biome, name) %>% 
  summarise(resid = mean(resid, na.rm = TRUE)) %>% 
  ungroup() %>% 
  drop_na()

pco2_product_biome_monthly_flux_attribution <-
  bind_rows(
    pco2_product_biome_monthly_flux_attribution,
    pco2_product_biome_monthly_flux_attribution_total
  )


pco2_product_biome_annual_flux_attribution <-
  pco2_product_biome_monthly_flux_attribution %>%
  group_by(year, biome, name) %>%
  summarise(resid = mean(resid)) %>%
  ungroup()


pco2_product_biome_monthly_flux_attribution %>%
  filter(year == {{year_anom}},
         biome %in% c("Global", key_biomes)) %>%
  ggplot() +
  geom_hline(yintercept = 0) +
  geom_path(
    aes(month, resid)
  ) +
  geom_point(
    aes(month, resid),
    shape = 21,
    alpha = 0.5,
    col = "grey30"
  ) +
  scale_y_continuous(breaks = seq(-10,10,0.2)) +
  scale_x_continuous(position = "top", breaks = seq(1,12,3)) +
  labs(y = labels_breaks(unique("fgco2"))$i_legend_title) +
  facet_grid(biome ~ name,
             labeller = labeller(name = x_axis_labels),
             scales = "free_y",
             space = "free_y", 
             switch = "x") +
  theme(
    legend.title = element_blank(),
    axis.title.y = element_markdown(),
    strip.text.x.bottom = element_markdown(),
    strip.placement = "outside",
    strip.background.x = element_blank(),
    legend.position = "top"
  )


```

```{r flux_attribution_seasonal}

pco2_product_biome_monthly_flux_attribution %>%
  filter(biome %in% c("Global", key_biomes)) %>%
  p_season(title  = paste("Anomalies from predicted monthly mean"))


```

### Annual

```{r flux_attribution_annual, fig.asp=1}


pco2_product_biome_annual_flux_attribution <-
  full_join(
    pco2_product_biome_annual_flux_attribution %>%
      filter(year == {{year_anom}}) %>%
      select(-year),
    pco2_product_biome_annual_flux_attribution %>%
      filter(year != {{year_anom}}) %>%
      group_by(biome, name) %>%
      summarise(resid_mean = mean(abs(resid))) %>%
      ungroup()
  )

pco2_product_biome_annual_flux_attribution %>%
  filter(biome %in% c("Global", key_biomes)) %>% 
  mutate(product = {{product_name}}) %>% 
  group_split(product) %>%
  # head(1) %>%
  map(
    ~ ggplot(data = .x) +
      geom_col(aes("x", resid),
               position = "dodge2") +
      geom_col(
        aes(
          "x",
          resid_mean * sign(resid),
          col = paste0("Mean\nexcl.",{{year_anom}})
        ),
        position = "dodge2",
        fill = "transparent"
      ) +
      labs(y = labels_breaks(unique("fgco2"))$i_legend_title,
           title = .x$biome) +
      facet_grid(
        biome~name,
        labeller = labeller(name = x_axis_labels),
        switch = "x"
      ) +
      scale_color_grey() +
      theme(
        legend.title = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_markdown(),
        strip.text.x.bottom = element_markdown(),
        strip.placement = "outside",
        strip.background.x = element_blank(),
        legend.position = "top"
      )
  )


```

```{r write_fco2_decomposition_flux_attribution, results = FALSE}

pco2_product_biome_annual_flux_attribution %>%
  write_csv(
    paste0(
      "../data/",
      "{{product_name}}",
      "_",
      "{{year_anom}}",
      "_biome_annual_flux_attribution.csv"
    )
  )

pco2_product_biome_monthly_flux_attribution %>%
  write_csv(
    paste0(
      "../data/",
      "{{product_name}}",
      "_",
      "{{year_anom}}",
      "_biome_monthly_flux_attribution.csv"
    )
  )

pco2_product_biome_annual_fCO2_decomposition %>%
  write_csv(
    paste0(
      "../data/",
      "{{product_name}}",
      "_",
      "{{year_anom}}",
      "_biome_annual_fCO2_decomposition.csv"
    )
  )

pco2_product_biome_monthly_fCO2_decomposition %>%
  write_csv(
    paste0(
      "../data/",
      "{{product_name}}",
      "_",
      "{{year_anom}}",
      "_biome_monthly_fCO2_decomposition.csv"
    )
  )

rm(
  pco2_product_biome_annual_flux_attribution,
  pco2_product_biome_monthly_flux_attribution,
  pco2_product_biome_annual_fCO2_decomposition,
  pco2_product_biome_monthly_fCO2_decomposition
)

gc()

```


# Zonal mean sections

The following analysis is available for GOBMs only.

## Annual means

### {{year_anom}} anomaly

```{r {{year_anom}}_anomaly_zonal_mean_sections_annual, fig.asp=1, eval=model, results = FALSE}

pco2_product_zonal_mean_annual <-   pco2_product_zonal_mean %>%
  pivot_longer(-c(region, depth, lat, time, year, month)) %>%
  group_by(region, lat, depth, year, name) %>%
  summarise(value = mean(value)) %>%
  ungroup() %>%
  drop_na() %>%
  mutate(region = str_to_title(region))

pco2_product_zonal_mean_annual_anomaly <-
  pco2_product_zonal_mean_annual %>% 
  anomaly_determination(region, lat, depth)

pco2_product_zonal_mean_annual_anomaly %>%
  filter(year == {{year_anom}}) %>%
  group_split(name) %>%
  # head(3) %>%
  map(
    ~ ggplot(data = .x) +
      geom_contour_filled(aes(lat, depth, z = resid)) +
      scale_fill_discrete_divergingx(name = labels_breaks(.x %>% distinct(name))$i_legend_title) +
      guides(fill = guide_colorsteps(
        barheight = unit(8, "cm"),
        show.limits = TRUE
      )) +
      scale_y_continuous(trans = trans_reverser("sqrt"),
                         breaks = c(50,100,200,400)) +
      scale_x_continuous(breaks = seq(-100, 100, 20)) +
      coord_cartesian(expand = 0) +
      facet_wrap( ~ region, ncol = 1) +
      labs(y = "Depth (m)") +
      theme(legend.title = element_markdown())
  )

```


```{r write_anomaly_anomaly_year_zonal_mean_sections_annual, eval=model}

pco2_product_zonal_mean_annual_anomaly %>%
  write_csv(
    paste0(
      "../data/",
      "{{product_name}}",
      "_",
      "{{year_anom}}",
      "_zonal_mean_sections_annual.csv"
    )
  )

```

# Biome profiles

The following analysis is available for GOBMs only.

## Annual means

### {{year_anom}} anomaly

```{r {{year_anom}}_anomaly_profiles_annual, fig.asp=1, eval=model, results = FALSE}

pco2_product_profiles_annual <-   pco2_product_profiles %>%
  pivot_longer(-c(biome, depth, time, year, month)) %>%
  group_by(biome, depth, year, name) %>%
  summarise(value = mean(value)) %>%
  ungroup() %>%
  drop_na()

pco2_product_profiles_annual_anomaly <-
  pco2_product_profiles_annual %>% 
  anomaly_determination(biome, depth)

pco2_product_profiles_annual_anomaly %>%
  group_split(name) %>%
  # head(1) %>%
  map(
    ~ ggplot(data = .x) +
      geom_path(aes(resid, depth, group = year), col = "grey30", alpha = 0.5) +
      geom_path(data = .x %>% filter(year == {{year_anom}}),
                aes(resid, depth, col = as.factor(year)),
                linewidth = 1) +
      scale_y_continuous(trans = trans_reverser("sqrt"),
                         breaks = c(50,100,200,400)) +
      facet_wrap( ~ biome) +
      labs(y = "Depth (m)",
           x = labels_breaks(.x %>% distinct(name))$i_legend_title) +
      theme(legend.title = element_blank(),
            axis.title.x = element_markdown())
  )

```


```{r write_anomaly_year_profiles_annual, eval=model}

pco2_product_profiles_annual_anomaly %>%
  write_csv(
    paste0(
      "../data/",
      "{{product_name}}",
      "_",
      "{{year_anom}}",
      "_profiles_annual.csv"
    )
  )

```

## Monthly means

### {{year_anom}} anomaly

```{r {{year_anom}}_anomaly_profiles_monthly, fig.asp=1, eval=model, results = FALSE}

pco2_product_profiles_monthly <-   pco2_product_profiles %>%
  pivot_longer(-c(biome, depth, time, year, month)) %>%
  group_by(biome, depth, year, month, name) %>%
  summarise(value = mean(value)) %>%
  ungroup() %>%
  drop_na()

pco2_product_profiles_monthly_anomaly <-
  pco2_product_profiles_monthly %>% 
  anomaly_determination(biome, depth, month)

pco2_product_profiles_monthly_anomaly %>%
  filter(year == {{year_anom}}) %>% 
  group_split(name) %>%
  # head(1) %>%
  map(
    ~ ggplot(data = .x) +
      geom_vline(xintercept = 0) +
      geom_path(aes(resid, depth, col = as.factor(month)),
                linewidth = 1) +
      scale_color_scico_d(palette = "hawaii") +
      scale_y_continuous(trans = trans_reverser("sqrt"),
                         breaks = c(50, 100, 200, 400)) +
      facet_wrap(~ biome,
                 scales = "free_x") +
      labs(y = "Depth (m)",
           x = labels_breaks(.x %>% distinct(name))$i_legend_title) +
      theme(legend.title = element_blank(),
            axis.title.x = element_markdown())
  )

```


```{r write_anomaly_year_profiles_monthly, eval=model}

pco2_product_profiles_monthly_anomaly %>%
  write_csv(
    paste0(
      "../data/",
      "{{product_name}}",
      "_",
      "{{year_anom}}",
      "_profiles_monthly.csv"
    )
  )

```
