---
title: "Biome mask"
author: "Jens Daniel Müller"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r parent, child = "/nfs/kryo/work/jenmueller/emlr_cant/utilities/setup.Rmd"}
# this chunk runs the code stored in setup.Rmd
# if required, please refer to instructions given here:
# https://jdblischak.github.io/workflowr/articles/wflow-07-common-code.html
```

```{r load_libraries_specific, include = FALSE}

library(stars)
library(khroma)

```

# Read data

In this study, we use the biome mask from RECCAP2, a modification of the original mask developed by Fay and McKinley (2014). 

```{r define_reccap_path}

path_reccap2 <-
  "/nfs/kryo/work/datasets/gridded/ocean/interior/reccap2/"

```


```{r read_region_mask}

print("RECCAP2_region_masks_all_v20221025.nc")

biome_mask <-
  read_ncdf(
    paste(
      path_reccap2,
      "supplementary/RECCAP2_region_masks_all_v20221025.nc",
      sep = ""
    )
  ) %>%
  as_tibble()


biome_mask <-
  biome_mask %>%
  mutate(lon = if_else(lon < 20, lon + 360, lon))


```

# Basemap and landmask

```{r basemap}

land_mask <- biome_mask %>%
  filter(seamask == 0) %>% 
  select(lon, lat)

map <- ggplot(land_mask,
              aes(lon, lat)) +
  geom_tile(fill = "grey80") +
  scale_y_continuous(breaks = seq(-60,60,30)) +
  scale_x_continuous(breaks = seq(0,360,60)) +
  coord_quickmap(expand = 0, ylim = c(-60, 80)) +
  theme(axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank())

map

map %>%
  write_rds(paste0("../data/","map.rds"))

```

# Global ocean

Global ocean estimates are integrated or averaged across the following areas.

```{r global_ocean_regions, fig.asp=0.5}

biome_mask <- biome_mask %>%
  filter(seamask == 1) %>% 
  select(lon, lat, atlantic:southern) %>% 
  pivot_longer(atlantic:southern,
               names_to = "region",
               values_to = "biome") %>%
  mutate(biome = as.character(biome))

biome_mask <- biome_mask %>%
  filter(biome != "0")

biome_mask <- biome_mask %>%
  mutate(biome = paste(region, biome, sep = "_"))

biome_mask <- biome_mask  %>% 
  mutate(biome = case_when(
    biome == "atlantic_1" ~ "NA-SPSS",
    biome == "atlantic_2" ~ "NA-STSS",
    biome == "atlantic_3" ~ "NA-STPS",
    biome == "atlantic_4" ~ "AEQU",
    biome == "atlantic_5" ~ "SA-STPS",
    # biome == "atlantic_6" ~ "MED",
    biome == "pacific_1" ~ "NP-SPSS",
    biome == "pacific_2" ~ "NP-STSS",
    biome == "pacific_3" ~ "NP-STPS",
    biome == "pacific_4" ~ "PEQU-W",
    biome == "pacific_5" ~ "PEQU-E",
    biome == "pacific_6" ~ "SP-STSS",
    biome == "indian_1" ~ "Arabian Sea",
    biome == "indian_2" ~ "Bay of Bengal",
    biome == "indian_3" ~ "Equatorial Indian",
    biome == "indian_4" ~ "Southern Indian",
    # biome == "arctic_1" ~ "ARCTIC-ICE",
    # biome == "arctic_2" ~ "NP-ICE",
    # biome == "arctic_3" ~ "NA-ICE",
    # biome == "arctic_4" ~ "Barents",
    # str_detect(biome, "arctic") ~ "Arctic",
    # biome == "southern_1" ~ "SO-STSS",
    # biome == "southern_2" ~ "SO-SPSS",
    # biome == "southern_3" ~ "SO-ICE",
    TRUE ~ "other"
  ))

biome_mask <-
  biome_mask %>%
  filter(biome != "other")

map +
  geom_tile(data = biome_mask,
            aes(lon, lat, fill = region)) +
  labs(title = "Considered ocean regions") +
  scale_fill_muted() +
  theme(legend.title = element_blank())

```


# Biomes

## All biomes

In the following, all individual biomes are plotted by ocean region.

```{r all_biomes, fig.asp=0.5}

biome_mask %>%
  group_split(region) %>%
  # head(1) %>%
  map( ~ map +
         geom_tile(data = .x,
                   aes(lon, lat, fill = biome)) +
         labs(title = paste("Region:", .x$region)) +
         scale_fill_okabeito())

```

## Key biomes

Following key biomes are highlighted throughout the analysis.

```{r key_biomes, fig.asp=0.5}

key_biomes <- c("NA-SPSS",
                "NA-STPS",
                "NP-SPSS",
                "PEQU-E")

key_biomes %>% 
  write_rds("../data/key_biomes.rds")

map +
  geom_tile(data = biome_mask %>% filter(biome %in% key_biomes),
            aes(lon, lat, fill = biome)) +
  labs(title = "Selected biomes to highlight") +
  scale_fill_muted() +
  theme(legend.title = element_blank())


biome_mask %>%
  select(-biome) %>% 
  write_rds("../data/region_mask.rds")

biome_mask <-
  biome_mask %>%
  select(-region)

biome_mask %>%
  write_rds("../data/biome_mask.rds")

```

## Super biomes

In addition to biomes, we focus our analysis on following combined super biomes.

```{r super_biomes, fig.asp=0.5}


super_biome_mask <- biome_mask  %>%
  mutate(
    biome = case_when(
      str_detect(biome, "NA-") ~ "North Atlantic",
      str_detect(biome, "NP-") ~ "North Pacific",
      # str_detect(biome, "SO-") ~ "Southern Ocean",
      TRUE ~ "other"
    )
  )

super_biome_mask <-
  super_biome_mask %>%
  filter(biome != "other")

map +
  geom_tile(data = super_biome_mask,
            aes(lon, lat, fill = biome)) +
  labs(title = "Selected super biomes") +
  scale_fill_muted() +
  theme(legend.title = element_blank())

super_biomes <-
  super_biome_mask %>%
  distinct(biome) %>%
  pull()

super_biomes %>%
  write_rds("../data/super_biomes.rds")

super_biome_mask %>%
  write_rds("../data/super_biome_mask.rds")

```
